---
title: "climate_rodent_analysis"
author: "R Pienaar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load libraries

```{r}

install.packages("DHARMa")
install.packages("performance")

library(DHARMa)
library(performance)
library(tidyverse)
library(ggplot2)
library(vegan)
library(lubridate)
library(caret)
library(ggpubr)
library(MuMIn)

```

Input data

```{r}

rodent_data <- read_csv("C:/Users/ryanp/Documents/Dissertation/Raw data/Mojave/Lytle_rodents/Summary_all.csv")
clim_data <- read.csv("C:/Users/ryanp/Documents/Dissertation/Chapter2_Rodents/Rodent_analysis/lytle_clim.csv")

```

Functions

```{r}

calc_rich <- function(data, col_start, col_end) {
  columns <- c(col_start:col_end)
  subset_data <- data[, columns]
  positive_columns_per_row <- rowSums(subset_data > 0)
  return(positive_columns_per_row)
}

calc_abun <- function(data, col_start, col_end) {
  columns <- c(col_start:col_end)
  subset_data <- data[, columns]
  total <- rowSums(subset_data)
  return(total)
}

simpsons_diversity <- function(data, col_start, col_end) {
  counts <- as.numeric(row[col_start:col_end])
  counts <- counts[!is.na(counts)]
  n <- sum(counts)
  if (n == 0) return(0)
  proportions <- counts / n
  return(1 - sum(proportions^2))
}

```



organize data
```{r}


#head(rodent_data)
rodent_data <- rodent_data |> replace(is.na(rodent_data),0)
#summary(rodent_data)

rodent_data <- rodent_data |> 
  mutate(Date = as.Date(paste("01",Date, sep = "-"), format = "%d-%b-%y"))

clim_data <- clim_data |> 
  mutate(Date = as.Date(paste("01",Date, sep = "-"), format = "%d-%b-%y"))

rodent_data$abundance <- calc_abun(rodent_data,5,20)
rodent_data$richness <- calc_rich(rodent_data,5,20)
rodent_data$simp_div <- diversity(rodent_data[,5:20], index = "simpson")
rodent_data$shan_div <- diversity(rodent_data[,5:20], index = "shannon")

scale_clim <- scale(clim_data[,3:9])

clim_data[,3:9] <- scale_clim

#head(clim_data)
#head(data)
data <- merge(rodent_data, clim_data, by = "Date", all.x = TRUE)

# Create lagged date in climate data
clim_data <- clim_data %>%
  mutate(Lagged_Date = Date %m-% years(1))

lagged_clim_data <- clim_data %>%
  select(Lagged_Date, ppt..mm., tmin..degrees.C., tmean..degrees.C., tmax..degrees.C., tdmean..degrees.C., vpdmin..hPa., vpdmax..hPa.) %>%
  rename(
    lag.ppt = ppt..mm.,
    lag.tmin..degrees.C. = tmin..degrees.C.,
    lag.tmean = tmean..degrees.C.,
    lag.tmax..degrees.C. = tmax..degrees.C.,
    lag.tdmean = tdmean..degrees.C.,
    lag.vmin = vpdmin..hPa.,
    lag.vmax = vpdmax..hPa.
  )

# Merge original data with the climate data
data <- data %>%
  left_join(lagged_clim_data, by = c("Date"= "Lagged_Date"))

names <- c("ppt", "tmin", "tmean", "tmax", "tdmean", "vmin", "vmax", "month", "season", "yst", 
           'lag.ppt', "lag.tmin", "lag.tmean", "lag.tmax", "lag.tdmean", "lag.vmin", "lag.vmax")

colnames(data)[26:42]<- names

data$season <- as.factor(data$season)
data$Fire <- as.factor(data$Fire)
data$BLOCK <- as.factor(data$BLOCK)

```


Explore

```{r, fig.height=8, fig.width=10}

# Total rodents Across all Plots
ggplot(data = data, aes(x=Date, y=abundance))+
  geom_smooth()+
  theme_classic()

# Compare rodents access to rodent exclusion over time
ggplot(data = data, aes(x=Date, y=abundance, color = Rodents))+
  geom_smooth()+
  theme_classic()

# Compare burned to unburned over time but we're only going to use rodent access from here
data <- data[data$Rodents=="S",]

# Question 1: Rodent abundance burned vs unburned
plotq1_1 <- ggplot(data = data, aes(x=Date, y=abundance, color = Fire))+
  geom_smooth()+
  theme_classic()+
  xlab("Year")+
  ylab("Abundance")+
  scale_color_manual(values = c("U" = "green4", "B"= "black"),label = c("Burned", 
                            "Unburned"))+
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text   = element_text(size=12, family = "serif", colour = "black"),
        legend.title  = element_text(size=12, family = "serif", colour = "black"),
        panel.border  = element_blank())



# Question 1: Rodent species richness burned vs unburned

plotq1_2 <- ggplot(data = data, aes(x=Date, y=richness, color = Fire))+
  geom_smooth()+
  theme_classic()+
  xlab("Year")+
  ylab("Species Richness")+
  scale_color_manual(values = c("U" = "green4", "B"= "black"),label = c("Burned", 
                            "Unburned"))+
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text   = element_text(size=12, family = "serif", colour = "black"),
        legend.title  = element_text(size=12, family = "serif", colour = "black"),
        panel.border  = element_blank())

# Question 1: Simpsons diversity burned vs unburned

plotq1_3 <- ggplot(data = data, aes(x=Date, y=simp_div, color = Fire))+
  geom_smooth()+
  theme_classic()+
  xlab("Year")+
  ylab("Simpsons Diversity")+
  scale_color_manual(values = c("U" = "green4", "B"= "black"),label = c("Burned", 
                            "Unburned"))+
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text   = element_text(size=12, family = "serif", colour = "black"),
        legend.title  = element_text(size=12, family = "serif", colour = "black"),
        panel.border  = element_blank())

# Question 1: Shannons diversity burned vs unburned

plotq1_4 <- ggplot(data = data, aes(x=Date, y=shan_div, color = Fire))+
  geom_smooth()+
  theme_classic()+
  xlab("Year")+
  ylab("Shannons Diversity")+
  scale_color_manual(values = c("U" = "green4", "B"= "black"),label = c("Burned", 
                            "Unburned"))+
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black"),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text   = element_text(size=12, family = "serif", colour = "black"),
        legend.title  = element_text(size=12, family = "serif", colour = "black"),
        panel.border  = element_blank())

q1plot <- ggarrange(plotq1_1,plotq1_2,plotq1_3,plotq1_4,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2,
                    common.legend = TRUE,
                    legend = "right",
                    font.label=list(color="black",size=12,family = "serif"))

q1plot

```
Do an ANOVA for question 1 to determine if there is a difference between burned and unburned areas

```{r}

head(data)

str(data)
data$BLOCK <- as.numeric(data$BLOCK)
ano.abun <- aov(data = data, abundance~Fire*Date + (1|BLOCK))
summary(ano.abun)

ano.rich <- aov(data = data, richness~ Fire*Date + (1|BLOCK))
summary(ano.rich)

ano.si <- aov(data = data, simp_div~ Fire*Date + (1|BLOCK))
summary(ano.si)

ano.sh <- aov(data = data, shan_div ~ Fire*Date + (1|BLOCK))
summary(ano.sh)


```


Start the analysis

Check for correlation in the data

```{r}

clim_matrix <- as.matrix(data[seq(from =1, to =301, by = 10),26:32])
clim_matrix <- na.omit(clim_matrix)
cor(clim_matrix)

# Keep tmean because all temps are correlated, keep vpdmax over vpdmin



```
Reduce Climate covariates to tmean, tdmean and ppt

Add the previous samples abundance as a covariate

```{r}

data <- data |>
  arrange(BLOCK, Fire, Date) |>
  group_by(BLOCK, Fire) |>
  mutate(lag.abun = lag(abundance))

data[,c(1,21,43)]

```


Use different types of models
First will be a linear mixed effects model

```{r}

library(lmerTest)
library(lme4)

### Split training and testing data

set.seed(123)
split <- initial_split(data, prop = 0.75)

train <- training(split)
train <- na.omit(train)
test <- testing(split)
test <- na.omit(test)

```

Direct comparison between burned and unburned plots

```{r}

ggplot(data = data, aes(x= Fire, y = abundance))+
 stat_summary(geom = "bar", fun = mean, position = "dodge", width = 0.3)+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,12))+
  coord_cartesian(ylim=c(0, 4))+
  ylab("Abundance")

ggplot(data = data, aes(x= Fire, y = richness))+
 stat_summary(geom = "bar", fun = mean, position = "dodge", width = 0.3)+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,6))+
  coord_cartesian(ylim=c(0, 2)) +
  ylab("Species Richness")

ggplot(data = data, aes(x= Fire, y = simp_div))+
 stat_summary(geom = "bar", fun = mean, position = "dodge", width = 0.3)+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,6))+
  coord_cartesian(ylim=c(0, 1))+
  ylab("Simpsons Diversity")

ggplot(data = data, aes(x= Fire, y = shan_div))+
 stat_summary(geom = "bar", fun = mean, position = "dodge", width = 0.3)+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.1, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,6))+
  coord_cartesian(ylim=c(0, 1))+
  ylab("Shannons Diversity")

```
Break it up by response variable

#Abundance GLMS

```{r}

ggplot(data = train, aes(y = abundance, x = yst, colour = Fire))+
  geom_smooth()+
  theme_classic()

ggplot(data = na.omit(train), aes(y = abundance, x = season, fill = Fire))+
  stat_summary(geom = "bar", fun = mean, position = "dodge")+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,12))+
  coord_cartesian(ylim=c(0, 5))+
  ylab("Abundance")

```


Information Theoretic Approach

Create 10 Models with less than 10 covariates each with a hypothesis as to why that would be a an appropriate model


```{r}

library(lme4)

data$BLOCK <- as.factor(data$BLOCK)

###

abun.null <- glmer(data = train, abundance ~ 1 +(1|BLOCK/yst), family = "poisson") # Null Model

abun.global <- glmer(data = train, abundance ~ ppt + lag.ppt + tmean + lag.tmean
                     + tdmean + lag.tdmean + Fire + (1|BLOCK/yst), family = "poisson") # Hyp: All factors contribute to rodent abundance

abun.fire <- glmer(data = train, abundance ~ Fire + (1|BLOCK/yst), family = "poisson") # Hyp: Fire is the main driver of rodent abundance through vegetation removal

abun.current <- glmer(data = train, abundance ~ ppt + tmean + tdmean + (1|BLOCK/yst),
                      family = "poisson")                                             # Hyp: Current weather conditions drive rodent abundance

abun.lag <- glmer(data = train, abundance ~ lag.ppt + lag.tmean + lag.tdmean + (1|BLOCK/yst),
                      family = "poisson")                                             # Hyp: Previous years weather drives rodent abundance

abun.ppt <- glmer(data = train, abundance ~ lag.ppt + ppt  + (1|BLOCK/yst),
                      family = "poisson")                                             # Hyp: Precipitation drives rodent abundance

abun.temp <- glmer(data = train, abundance ~ tmean + lag.tmean + tdmean + lag.tdmean +
                     (1|BLOCK/yst), family = "poisson")                               # Hyp: Temperature drives rodent abundance

abun.fire.ppt <- glmer(data = train, abundance ~ ppt * Fire + lag.ppt * Fire +
                         (1|BLOCK/yst), family = "poisson")                               # Hyp: Interaction between fire and precipitation drives rodent abundance

abun.fire.temp <- glmer(data = train, abundance ~ Fire * tmean + Fire * tdmean +
                          Fire * lag.tmean + Fire * lag.tdmean +(1|BLOCK/yst),
                        family = "poisson")                                         # Hyp: The interaction between fire and temperature drives rodent abundance

abun.sel <- model.sel(abun.null, abun.global, abun.fire, abun.current, abun.lag,
                      abun.ppt, abun.temp, abun.fire.ppt, abun.fire.temp)

abun.sel

```

Validate model

```{r}


testDispersion(abun.lag)

testZeroInflation(abun.lag)

performance(abun.lag)

simulateResiduals(abun.lag, plot = TRUE)

```

Validate best model

```{r}

par(mfrow=c(2,1))
qqnorm(resid(abun.lag), main="normal qq-plot, residuals")
qqline(resid(abun.lag))

#qqnorm(ranef(abun8)$id[,1])
#qqline(ranef(abun8)$id[,1])

plot(fitted(abun.lag), resid(abun.lag)) #residuals vs fitted
abline(h=0)


#dat_kackle$fitted <- fitted(abun8)    #fitted vs observed
#plot(your_data$fitted, jitter(your_data$total,0.1))
#abline(0,1)

```

Cross validate best model

```{r}

abun.lag.pred <- predict(abun.lag, test)

plot(test$abundance, abun.lag.pred)

summary(lm(test$abundance~ abun.lag.pred))

```



## Abundance Random Forest

```{r}

library(randomForest)
library(pROC)

rfdata <- na.omit(train)
rftest <- na.omit(test)
head(rfdata)
rfdata[1:6,-c(5:20)]

set.seed(123)

abun.rf <- randomForest(data = rfdata, abundance ~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst )

abun.rf

# Version 2

abun.rf2 <- randomForest(data = rfdata, abundance ~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst +lag.abun)
abun.rf2

plot(abun.rf)

varImpPlot(abun.rf)

varimp <- data.frame(abun.rf$importance)

varimp$variable <- rownames(varimp)

varimp <- varimp[order(varimp$IncNodePurity),]

head(varimp)

ggplot(data= varimp, aes(x= fct_reorder(variable, IncNodePurity), y = IncNodePurity))+
  geom_col()+
  theme_classic()+
  ylab("Relative variable importance")+
  xlab("Variable")+
  theme(plot.title    = element_text(size=12, family = "serif", colour = "black"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12, family = "serif", colour = "black"),
        axis.title.y  = element_text(size=12, family = "serif", colour = "black"),
        axis.text.x   = element_text(size=12, family = "serif", colour = "black", angle = 45, hjust = 1),
        axis.text.y   = element_text(size=12, family = "serif", colour = "black"),
        legend.text   = element_text(size=12, family = "serif", colour = "black"),
        legend.title  = element_text(size=12, family = "serif", colour = "black"),
        panel.border  = element_blank())



abun.rf.pred <- predict(abun.rf, newdata = rftest)

plot(abun.rf.pred, test$abundance)
summary(lm(abun.rf.pred~ test$abundance))

roc.rf.abun <- roc(rftest$abundance, abun.rf.pred)

auc(roc.rf.abun)

#### Repeat RF but with k-folds cross validation ####

custom_control <- trainControl(
  method = "cv",       # Cross-validation
  number = 5,          # Number of folds
  verboseIter = TRUE,  # Output the progress
  savePredictions = "final"
)

set.seed(123)  # Set seed for reproducibility
rf_abun_model <- train(
  abundance~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst,            # Formula for the model
  data = rfdata,            # Data to be used
  method = "cforest",   # SVM with radial basis function kernel
  trControl = custom_control,
  tuneLength = 10         # Tune over 10 different parameter values
)
print(rf_abun_model)

typeof(abun.rf)

typeof(rf_abun_model)


```

Try the random forest but with seasonal precipitation

```{r}

season_rf <- rfdata
summary(season_rf$season)
season_rf <-season_rf %>%
  mutate(season = factor(season, levels = c("Spring", "Summer", "Winter"))) %>%  # Ensure correct order
  pivot_wider(names_from = season, values_from = lag.ppt, names_prefix = "lag.ppt_")

```



## Abundance SVM

Get an abundance SVM classification, see what that looks like
```{r}


library(e1071)



### Single SVM ###
svm.abun <- svm(data=  rfdata, abundance~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst, kernel = "radial", cost = 5)

svm.abun

svm.abun.pred <- predict(svm.abun, newdata = rftest)

plot(svm.abun.pred, rftest$abundance)

summary(lm(svm.abun.pred~ rftest$abundance))
sqrt(mean((svm.abun.pred - rftest$abundance)^2))

roc.svm.abun <- roc(rftest$abundance, svm.abun.pred)

auc(roc.svm.abun)


# Calculate R-squared (R²)
ss_total <- sum((rftest$abundance - mean(rftest$abundance))^2)
ss_res <- sum((svm.abun.pred - rftest$abundance)^2)
r_squared <- 1 - (ss_res / ss_total)
print(paste("R-squared:", r_squared))

### why not just lm for r sqaured
summary(lm(rftest$abundance~svm.abun.pred))


#### Repeat SVM but with k-folds cross validation ####

custom_control <- trainControl(
  method = "cv",       # Cross-validation
  number = 5,          # Number of folds
  verboseIter = TRUE,  # Output the progress
  savePredictions = "final"
)

set.seed(123)  # Set seed for reproducibility
svm_abun_model <- train(
  abundance~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst,            # Formula for the model
  data = rfdata,            # Data to be used
  method = "svmRadial",   # SVM with radial basis function kernel
  trControl = custom_control,
  tuneLength = 10         # Tune over 10 different parameter values
)
print(svm_abun_model)




```

## Abundance K-Nearest neighbor

```{r}

library(class)

knn_abun_model <- knnreg(abundance~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst,
                         k=1,
                         data = rfdata)


summary(knn_abun_model)

knn_abun_pred <- predict(knn_abun_model, rftest)
knn_abun_pred

plot(knn_abun_pred, rftest$abundance)
summary(lm(knn_abun_pred~ rftest$abundance))


# No kfold cross validation method??

```
## Abundance NN

```{r, warning=FALSE}

library(neuralnet)

rfdata_dummy <- model.matrix(~ Fire + tmin + tmean + tmax + tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax + lag.tdmean + lag.vmax + lag.vmin + season + yst, data = rfdata)

rfdata_dummy <- data.frame(abundance = rfdata$abundance, rfdata_dummy)

rftest_dummy <- model.matrix(~ Fire + tmin + tmean + tmax + tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax + lag.tdmean + lag.vmax + lag.vmin + season + yst, data = rftest)


nn_abun_model <- neuralnet(abundance~ FireU + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ seasonSummer+ seasonWinter + yst,
                          data = rfdata_dummy)

nn_abun_pred <- predict(nn_abun_model, rftest_dummy)

summary(lm(nn_abun_pred ~ rftest$abundance))

#### Repeat Neuaral Net but with k-folds cross validation ####

custom_control <- trainControl(
  method = "cv",       # Cross-validation
  number = 5,          # Number of folds
  verboseIter = TRUE,  # Output the progress
  savePredictions = "final"
)

set.seed(123)  # Set seed for reproducibility
nn_abun_model <- train(
  abundance~ FireU + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ seasonSummer+ seasonWinter + yst,# Formula for the model
  data = rfdata_dummy,            # Data to be used
  method = "neuralnet",   # 
  trControl = custom_control,
  tuneLength = 10         # Tune over 10 different parameter values
)
print(nn_abun_model)



```


## Abundance XGBoost

```{r}

options(warn = -1)

library(xgboost)
library(Ckmeans.1d.dp)
library(gbm)


rfdata_dummy <- model.matrix(~ Fire + tmin + tmean + tmax + tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax + lag.tdmean + lag.vmax + lag.vmin + season + yst, data = rfdata)

xgdata <- xgb.DMatrix(rfdata_dummy, label= rfdata$abundance)

rfdata_dummy <- data.frame(abundance = rfdata$abundance, rfdata_dummy)


xg_abun_model <- xgboost(data = xgdata, 
              objective="reg:squarederror", 
              nrounds = 50, 
              verbose = 0)


xgb.importance(model = xg_abun_model)
xgb.ggplot.importance(xgb.importance(model = xg_abun_model), measure = "Frequency")

#### Repeat Neural Net but with k-folds cross validation ####

custom_control <- trainControl(
  method = "cv",       # Cross-validation
  number = 5,          # Number of folds
  verboseIter = TRUE,  # Output the progress
  savePredictions = "final"
)

set.seed(123)  # Set seed for reproducibility
xg_abun_model <- train(
  abundance~ FireU + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ seasonSummer+ seasonWinter + yst,# Formula for the model
  data = rfdata_dummy,            # Data to be used
  method = "xgbTree",   # 
  trControl = custom_control,
  tuneLength = 10,# Tune over 10 different parameter values
  verbose = FALSE,
  verbosity = 0)
print(xg_abun_model)

```



#Richness

##Fire and Time

Lets start with a model of time and fire to see how they affect the 4 metrics

```{r}

rich.ft <- glmer(data = train, richness ~ Fire + yst + Fire * yst+(1|BLOCK), family = "poisson")
summary(rich.ft)

rich.fs <- glmer(data = train, richness ~ Fire + season + Fire * season+(1|BLOCK), family = "poisson")
summary(rich.fs)

rich.fp <- glmer(data = train, richness ~ Fire + ppt + Fire * ppt+(1|BLOCK), family = "poisson")
summary(rich.fp)

rich.ft <- glmer(data = train, richness ~ Fire + tmean + Fire * tmean+(1|BLOCK), family = "poisson")
summary(rich.ft)

rich.ftd <- glmer(data = train, richness ~ Fire + tdmean + Fire * tdmean+(1|BLOCK), family = "poisson")
summary(rich.ftd)

rich.fv <- glmer(data = train, richness ~ Fire + vmax + Fire * vmin+(1|BLOCK), family = "poisson")
summary(rich.fv)

rich.flp <- glmer(data = train, richness ~ Fire + lag.ppt + Fire * lag.ppt+(1|BLOCK), family = "poisson")
summary(rich.flp)

rich.flt <- glmer(data = train, richness ~ Fire + lag.tmean + Fire * lag.tmean+(1|BLOCK), family = "poisson")
summary(rich.flt)

rich.fltd <- glmer(data = train, richness ~ Fire + lag.tdmean + Fire * lag.tdmean+(1|BLOCK), family = "poisson")
summary(rich.fltd)

rich.flv <- glmer(data = train, richness ~ Fire + lag.vmin + Fire * lag.vmin+(1|BLOCK), family = "poisson")
summary(rich.flv)

rich.sel <- model.sel(rich.flp, rich.flt, rich.fltd, rich.flv, rich.fp, rich.fs, rich.ft, rich.ftd, rich.fv)
rich.sel


ggplot(data = train, aes(y = richness, x = yst, colour = Fire))+
  geom_smooth()+
  theme_classic()

ggplot(data = na.omit(train), aes(y = richness, x = season, fill = Fire))+
  stat_summary(geom = "bar", fun = mean, position = "dodge")+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,6))+
  coord_cartesian(ylim=c(0, 3))+
  ylab("Species Richness")


```



## Complex model selection

```{r}

rich1 <- glmer(data = train, richness ~ yst + lag.tdmean + Fire + vmin+(1|BLOCK), family = "poisson")
rich2 <- glmer(data = train, richness ~ yst + lag.tdmean + Fire + ppt+(1|BLOCK), family = "poisson")
rich3 <- glmer(data = train, richness ~ yst + tdmean + Fire + vmin+(1|BLOCK), family = "poisson")
rich4 <- glmer(data = train, richness ~ yst + lag.tdmean + Fire + lag.ppt+(1|BLOCK), family = "poisson")
rich5 <- glmer(data = train, richness ~ lag.ppt + lag.tdmean + Fire + vmin+(1|BLOCK), family = "poisson")
rich6 <- glmer(data = train, richness ~ ppt + yst + Fire+(1|BLOCK), family = "poisson")
rich7 <- glmer(data = train, richness ~ ppt + yst + Fire + lag.ppt+(1|BLOCK), family = "poisson")


rich.sel <- model.sel(rich.flp, rich.flt, rich.fltd, rich.flv, rich.fp, rich.fs, rich.ft, rich.ftd, rich.fv,
                      rich1, rich2, rich3, rich4, rich5, rich6, rich7)
rich.sel

abuntest <- glmmTMB(data = train, abundance ~ ppt + yst + Fire + lag.ppt + (1|BLOCK/yst), family = poisson)

table(train$BLOCK)
table(train$BLOCK, train$yst)
performance(abuntest)

summary(abuntest)

testDispersion(rich4)
testZeroInflation(rich4)

```
Evaluate linear models

```{r}

rich.pred <- predict(rich4,newdata = test)

summary(lm(rich.pred~test$richness))

```


## Richness Random Forest

```{r}

rich.rf <- randomForest(data = rfdata, richness ~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst)

rich.rf

plot(rich.rf)

varImpPlot(rich.rf)

rich.rf.pred <- predict(rich.rf, newdata = rftest)

roc.rf.rich <- roc(rftest$richness, rich.rf.pred)

auc(roc.rf.rich)

```

##Richness SVM


#Simpsons Diversty



Lets start with a model of time and fire to see how they affect the 4 metrics

```{r}


si.ft <- glmer(data = train, simp_div ~ Fire + yst + Fire * yst+(1|BLOCK))
summary(si.ft)

si.fs <- glmer(data = train, simp_div ~ Fire + season + Fire * season+(1|BLOCK))
summary(si.fs)

si.fp <- glmer(data = train, simp_div ~ Fire + ppt + Fire * ppt+(1|BLOCK))
summary(si.fp)

si.ft <- glmer(data = train, simp_div ~ Fire + tmean + Fire * tmean+(1|BLOCK))
summary(si.ft)

si.ftd <- glmer(data = train, simp_div ~ Fire + tdmean + Fire * tdmean+(1|BLOCK))
summary(si.ftd)

si.fv <- glmer(data = train, simp_div ~ Fire + vmax + Fire * vmin+(1|BLOCK))
summary(si.fv)

si.flp <- glmer(data = train, simp_div ~ Fire + lag.ppt + Fire * lag.ppt+(1|BLOCK))
summary(si.flp)

si.flt <- glmer(data = train, simp_div ~ Fire + lag.tmean + Fire * lag.tmean+(1|BLOCK))
summary(si.flt)

si.fltd <- glmer(data = train, simp_div ~ Fire + lag.tdmean + Fire * lag.tdmean+(1|BLOCK))
summary(si.fltd)

si.flv <- glmer(data = train, simp_div ~ Fire + lag.vmin + Fire * lag.vmin+(1|BLOCK))
summary(si.flv)

si.sel <- model.sel(si.flp, si.flt, si.fltd, si.flv, si.fp, si.fs, si.ft, si.ftd, si.fv)
si.sel



ggplot(data = train, aes(y = simp_div, x = yst, colour = Fire))+
  geom_smooth()+
  theme_classic()

ggplot(data = na.omit(train), aes(y = simp_div, x = season, fill = Fire))+
  stat_summary(geom = "bar", fun = mean, position = "dodge")+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,6))+
  coord_cartesian(ylim=c(0, 1))+
  ylab("Simpsons Diversity")


```



## Simple model comparison

```{r}

si.sel <- model.sel(si.flp, si.flt, si.fltd, si.flv, si.fp, si.fs, si.ft, si.ftd, si.fv)
si.sel

```

## Complex model selection

```{r}

si1 <- glmer(data = train, simp_div ~ yst + lag.tdmean + Fire + vmin + (1|BLOCK))
si2 <- glmer(data = train, simp_div ~ yst + lag.tdmean + Fire + ppt + (1|BLOCK))
si3 <- glmer(data = train, simp_div ~ yst + tdmean + Fire + vmin + (1|BLOCK))
si4 <- glmer(data = train, simp_div ~ yst + lag.tdmean + Fire + lag.ppt + (1|BLOCK))
si5 <- glmer(data = train, simp_div ~ lag.ppt + lag.tdmean + Fire + vmin + (1|BLOCK))
si6 <- glmer(data = train, simp_div ~ ppt + yst + Fire + (1|BLOCK))
si7 <- glmer(data = train, simp_div ~ ppt + yst + Fire + lag.ppt + (1|BLOCK))


si.sel <- model.sel(si.flp, si.flt, si.fltd, si.flv, si.fp, si.fs, si.ft, si.ftd, si.fv,
                      si1, si2, si3, si4, si5, si6, si7)
si.sel


```

Validate model

```{r}

si.pred <- predict(si.flp, newdata = test)

plot(si.pred, test$simp_div)

summary(lm(si.pred~test$simp_div))

```




## Simpsons diversity Random Forest

```{r}

si.rf <- randomForest(data = rfdata, simp_div ~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst)

si.rf

plot(si.rf)

varImpPlot(si.rf)

si.rf.pred <- predict(si.rf, newdata = rftest)

roc.rf.si <- roc(rftest$simp_div, si.rf.pred)

auc(roc.rf.rich)

summary(lm(si.rf.pred~test$simp_div))

```



#Shannons Diversity



Lets start with a model of time and fire to see how they affect the 4 metrics

```{r}

sh.ft <- glmer(data = train, shan_div ~ Fire + yst + Fire * yst+ (1|BLOCK))
summary(sh.ft)

sh.fs <- glmer(data = train, shan_div ~ Fire + season + Fire * season+ (1|BLOCK))
summary(sh.fs)

sh.fp <- glmer(data = train, shan_div ~ Fire + ppt + Fire * ppt+ (1|BLOCK))
summary(sh.fp)

sh.ft <- glmer(data = train, shan_div ~ Fire + tmean + Fire * tmean+ (1|BLOCK))
summary(sh.ft)

sh.ftd <- glmer(data = train, shan_div ~ Fire + tdmean + Fire * tdmean+ (1|BLOCK))
summary(sh.ftd)

sh.fv <- glmer(data = train, shan_div ~ Fire + vmax + Fire * vmax+ (1|BLOCK))
summary(sh.fv)

sh.flp <- glmer(data = train, shan_div ~ Fire + lag.ppt + Fire * lag.ppt+ (1|BLOCK))
summary(sh.flp)

sh.flt <- glmer(data = train, shan_div ~ Fire + lag.tmean + Fire * lag.tmean+ (1|BLOCK))
summary(abun.flt)

sh.fltd <- glmer(data = train, shan_div ~ Fire + lag.tdmean + Fire * lag.tdmean+ (1|BLOCK))
summary(abun.fltd)

sh.flv <- glmer(data = train, shan_div ~ Fire + lag.vmax + Fire * lag.vmax+ (1|BLOCK))
summary(abun.flv)


ggplot(data = train, aes(y = shan_div, x = yst, colour = Fire))+
  geom_smooth()+
  theme_classic()

ggplot(data = na.omit(train), aes(y = shan_div, x = season, fill = Fire))+
  stat_summary(geom = "bar", fun = mean, position = "dodge")+
  stat_summary(geom = "errorbar", fun.data = mean_se, width = 0.2, position = position_dodge(0.9))+
  theme_classic()+
 scale_y_continuous(expand = c(0, 0), limits = c(0,6))+
  coord_cartesian(ylim=c(0, 1))+
  ylab("Shannons Diversity")


```



# Simple model selectoin

```{r}

sh.sel <- model.sel(sh.flp, sh.flt, sh.fltd,sh.flv, sh.fp, sh.fs,sh.ft, sh.ftd, sh.fv)
sh.sel

```
## Complex model selection

```{r}

library(glmmTMB)

sh1 <- glmmTMB(data = train, shan_div ~ yst + lag.tdmean + Fire + vmin + (1|BLOCK),  
               zi =~  yst + lag.tdmean + Fire + vmin + (1|BLOCK), family = nbinom2)

sh2 <- glmmTMB(data = train, shan_div ~ yst + lag.tdmean + Fire + ppt + (1|BLOCK),
             zi =~  yst + lag.tdmean + Fire + ppt + (1|BLOCK), family = nbinom2)

sh3 <- glmmTMB(data = train, shan_div ~ yst + tdmean + Fire + vmin + (1|BLOCK),
             zi=~ yst + tdmean + Fire + vmin + (1|BLOCK), family = nbinom2)

sh4 <- glmmTMB(data = train, shan_div ~ yst + lag.tdmean + Fire + lag.ppt + (1|BLOCK),
             zi =~ yst + lag.tdmean + Fire + lag.ppt + (1|BLOCK), family = nbinom2)

sh5 <- glmmTMB(data = train, shan_div ~ lag.ppt + lag.tdmean + Fire + vmin + (1|BLOCK),
             zi=~ lag.ppt + lag.tdmean + Fire + vmin + (1|BLOCK), family = nbinom2)

sh6 <- glmmTMB(data = train, shan_div ~ ppt + yst + Fire + (1|BLOCK),
             zi =~ ppt + yst + Fire + (1|BLOCK), family = nbinom2)

sh7 <- glmmTMB(data = train, shan_div ~ ppt + yst + Fire + lag.ppt + (1|BLOCK),
             zi =~ ppt + yst + Fire + lag.ppt + (1|BLOCK), family = nbinom2)

sh.sel <- model.sel(sh.flp, sh.flt, sh.fltd,sh.flv, sh.fp, sh.fs,sh.ft, sh.ftd, sh.fv, sh1, sh2, sh3,
                    sh4, sh5, sh6, sh7)
sh.sel


```

Check r squared

```{r}

sh.pred <- predict(sh.flp, newdata = test)

plot(sh.pred, test$shan_div)

summary(lm(sh.pred~test$shan_div))

```


## Random Forest

```{r}

sh.rf <- randomForest(data = rfdata, shan_div ~ Fire + tmin + tmean + tmax +
                          tdmean + vmax + vmin + ppt + lag.ppt + lag.tmin + lag.tmean + lag.tmax +
                          lag.tdmean + lag.vmax + lag.vmin + lag.ppt+ season + yst )

sh.rf

plot(sh.rf)

varImpPlot(sh.rf)

sh.rf.pred <- predict(sh.rf, newdata = rftest)

roc.rf.sh <- roc(rftest$shan_div, sh.rf.pred)

auc(roc.rf.sh)

importance(sh.rf)

```

