---
title: "Desert Rodent Analysis"
author: "R Pienaar"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits = 3)

library(tidyverse)
library(vegan)
library(ggpubr)
library(lme4)
library(lmerTest)
library(broom)
library(purrr)
library(randomForest)
library(glmmTMB)
library(ggbeeswarm)
library(patchwork)
library(MuMIn)
library(performance)
library(DHARMa)
library(robustbase)
library(cowplot)
library(lavaan)
library(piecewiseSEM)


```

# 0. Functions

Functions

```{r}

# Function to calculate species richness
calc_rich <- function(data, col_start, col_end) {
  columns <- c(col_start:col_end)
  subset_data <- data[, columns]
  positive_columns_per_row <- rowSums(subset_data > 0)
  return(positive_columns_per_row)
}

# Function to calculate abundance
calc_abun <- function(data, col_start, col_end) {
  columns <- c(col_start:col_end)
  subset_data <- data[, columns]
  total <- rowSums(subset_data)
  return(total)
}

# Common theme to be used for figures
common_theme <- theme(plot.title    = element_text(size=12,   colour = "black", hjust = 0.5, face="bold", family = "serif"),
        plot.subtitle = element_text(size=12, family = "serif"),
        axis.title.x  = element_text(size=12,  colour = "black", family = "serif"),
        axis.title.y  = element_text(size=12,  colour = "black", family = "serif"),
        axis.text.x   = element_text(size=12,  colour = "black", family = "serif"),
        axis.text.y   = element_text(size=12,  colour = "black", family = "serif"),
        legend.text   = element_text(size=12,  colour = "black", family = "serif"),
        legend.title  = element_text(size=12,  colour = "black", family = "serif"),
        panel.border  = element_blank())

```



Project Description:

Rodent communities in western US deserts play an important ecological role in plant community assembly through granivory and herbivory. In many areas, rodent populations are being affected by ecological change due to human-altered climate and fire regimes. This study aimed to test the effects of fire and climate variability on rodent community abundance, richness, and diversity over a 13-year period in the Mojave and Great Basin Deserts.

Description of data:
There are 2 csv's, one for the Great Basin and the other in the Mojave. These data run from 2011 to 2024, rodents were trapped 3 times a year initially but that changed to eventually only once a year for the last 3 years. The initial collection were also performed over 3 trapped nights but at a later stage adjusted to 2 trap nights (2022 to 2024). Each sampling occasion has a row for both burned and unburned plots for each of the 5 blocks. The data shows the number of unique individuals caught in a plot per sampling occasion, the number of each species is recorded in the species code columns and the total in the abundance column. The number of species are in the richness column and then diversity is determined from the species counts. Climate data was calculated using PRISM daily climate data from 2009 to 2024 for the 2 locations at 800m resolution. The 2 climate variables used were total precipitation and mean daily temperature. These were then used to calculate monthly totals for ppt and averages for tmean. We then calculated TAP (Total Annual Precipitation) and MAT (Mean Annual Temperature) as the sum of and average of the previous 12 months leading up to the sampling event. lag_TAP and lag_MAT were then calculated as the previous 12 months before that. Seasons were then defined by taking the average or total for the current season of the sampling occasion but only including months that had fully past (i.e if sampling was in April spring precipitation was only the average for March but if was in June spring precipitation included March through May). In this way the current years seasons didn't always go back a full year if sampling was in the middle of a season but lagged seasons all included a full year.

# 0. Input data

```{r}

data_gb <- read.csv("Data/GreatBasin_Rodents_Formatted.csv")
data_moj <- read.csv("Data/Mojave_Rodents_Formatted.csv")

```

# 1. Fire comparison

## 1.1 Fire Statistics

```{r}

# Add a site column so we can create a combined figure
data_gb$site <- "Great Basin"
data_moj$site <- "Mojave"

# Combined Great Basin and Mojave dataframe
combo<- rbind(data_gb |> select(Year, PLOT, Block, Abun, Rich, Div, site), data_moj |> select(Year, PLOT, Block, Abun, Rich, Div, site))

factor(data_gb$PLOT)
data_gb$PLOT <- factor(data_gb$PLOT, levels = c("US", "BS"))
data_moj$PLOT <- factor(data_moj$PLOT, levels = c("US", "BS"))

# Prevent scientific notation
options(scipen = 999)

# Use non-parametric comparison to see if theres a significant difference between burned and unburned plots
wilcox.test(data = data_gb, Abun ~ PLOT)
wilcox.test(data = data_gb, Rich ~ PLOT)
wilcox.test(data = data_gb, Div ~ PLOT)
wilcox.test(data = data_moj, Abun ~ PLOT)
wilcox.test(data = data_moj, Rich ~ PLOT)
wilcox.test(data = data_moj, Div ~ PLOT)

# This does the exact samw thing but uses a loop to make a table, probably every unneccessary but the table gets made
vars <- c("Abun","Rich","Div")
datasets <- list(data_gb = data_gb, data_moj = data_moj)

summ <- bind_rows(lapply(names(datasets), function(dn){
  df <- datasets[[dn]]
  
  # normalize PLOT labels to lowercase and rename for convenience
  df <- df %>% mutate(plot_l = tolower(as.character(PLOT)))
  
  lapply(vars, function(v){
    x <- df %>% filter(!is.na(.data[[v]]))
    
    # Means for burned vs unburned
    mean_b <- mean(x[[v]][x$plot_l %in% c("burned","bs")], na.rm=TRUE)
    mean_u <- mean(x[[v]][x$plot_l %in% c("unburned","us")], na.rm=TRUE)
    
    pct   <- ifelse(is.na(mean_u) | mean_u == 0, NA, 100*(mean_b-mean_u)/mean_u)
    
    # Wilcoxon test
    wt <- tryCatch({
      broom::tidy(wilcox.test(x[[v]] ~ x$plot_l, exact=FALSE))
    }, error = function(e) tibble(statistic=NA_real_, p.value=NA_real_))
    
    tibble(dataset=dn, variable=v,
           mean_burned = mean_b, mean_unburned = mean_u,
           pct_diff = round(pct,1),
           W = as.numeric(wt$statistic), 
           p_value = as.numeric(wt$p.value))
    
  }) %>% bind_rows()
}))

summ$p_value <- round(summ$p_value, digits = 4)

print(summ)

write.csv(summ, "Tables/Fire_Rodent_Table.csv", row.names = FALSE)

```

## 1.2 Fire Plots

```{r}

# Plot rodent abundance in burned and unburned plots
p.com.abun <- ggplot(data = combo, aes(x=site, y = Abun, fill = PLOT))+
  stat_summary(geom = "bar", fun = "mean", position = position_dodge(0.9), color = "black", width = 0.75)+
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.3)+
  xlab("")+
  labs(fill = "")+
  ylab("Rodent Abundance")+
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0, 4.5))+
  scale_fill_manual(values = c("grey30","white"), labels = c("Burned", "Unburned"))+
  theme_classic()+
  common_theme+
  annotate("text", x = 1, y = 4, label = "***", size = 4, family = "serif")+ # p < 0.001
  annotate("text", x = 2, y = 4, label = "ns", size = 4, family = "serif")   # p = 0.2

# Plot species richness in burned and unburned plots
p.com.rich <- ggplot(data = combo, aes(x=site, y = Rich, fill = PLOT))+
  stat_summary(geom = "bar", fun = "mean", position = position_dodge(0.9), color = "black", width = 0.75)+
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.3)+
  xlab("")+
  labs(fill = "")+
  ylab("Species Richness")+
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0, 2))+
  scale_fill_manual(values = c("grey30","white"), labels = c("Burned", "Unburned"))+
  theme_classic()+
  common_theme+
  annotate("text", x = 1, y = 1.7, label = "***", size = 4, family = "serif")+ # p < 0.001
  annotate("text", x = 2, y = 1.7, label = "*", size = 4, family = "serif")  # p = 0.02

# Plot Shannon's Diversity in burned and unburned plots
p.com.div <- ggplot(data = combo, aes(x=site, y = Div, fill = PLOT))+
  stat_summary(geom = "bar", fun = "mean", position = position_dodge(0.9), color = "black", width = 0.75)+
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.3)+
  xlab("")+
  labs(fill = "")+
  ylab("Shannon's Diversity")+
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0, 0.5))+
  scale_fill_manual(values = c("gray30","white"), labels = c("Burned", "Unburned"))+
  theme_classic()+
  common_theme +
  annotate("text", x = 1, y = 0.4, label = "**", size = 4, family = "serif")+ # p = 0.009
  annotate("text", x = 2, y = 0.4, label = "*", size = 4, family = "serif") # p = 0.01

# Combined plot
fireeffes <- ggarrange(p.com.abun,
          p.com.rich,
          p.com.div,
          nrow = 3,
          common.legend = TRUE,
          legend = 'right')
fireeffes

ggsave("Plots/Fire_Rodents_Plot.png",fireeffes, dpi = 300, height = 6, width = 4, units = "in")

```





# 2. Fire effects over time

## 2.1 Changes in Rodent metrics over time stats

```{r}


data_gb$Fire <- as.factor(data_gb$Fire)
data_moj$Fire <- as.factor((data_moj$Fire))

data_gb$Fire <- relevel(data_gb$Fire, ref = "U")
data_moj$Fire <- relevel(data_moj$Fire, ref = "U")


# Create a time column instead of just a year
data_gb$Time <- data_gb$Year - 2011
data_moj$Time <- data_moj$Year - 2011

# Check for temporal autocorrelation
acf(data_gb$Abun)
acf(data_gb$Rich)
acf(data_gb$Div)
acf(data_moj$Abun)
acf(data_moj$Rich)
acf(data_moj$Div)
# No Temporal autocorrelation


# Test the interactive effects of fire and time on all 3 metrics
# To meet assumptions we use robust linear regression for abundance and richness
# we use a generalized linear mixed model for Shannon's Diversity
m.gb.abun.time <- lmrob(data = data_gb, Abun ~ Fire * Time)
m.gb.rich.time <- lmrob(data = data_gb, Rich ~ Fire * Time)
m.gb.shan.time <- glmmTMB(data = data_gb, Div ~ Fire * Time,
                          family = tweedie(link = "log"))

m.moj.abun.time <- lmrob(data = data_moj, Abun ~ Fire * Time)
m.moj.rich.time <- lmrob(data = data_moj, Rich ~ Fire * Time)
m.moj.shan.time <- glmmTMB(data = data_moj, Div ~ Fire * Time,
                          family = tweedie(link = "log"))



models <- list(
  GB_Abundance = m.gb.abun.time,
  GB_Richness  = m.gb.rich.time,
  GB_Diversity = m.gb.shan.time,
  Moj_Abundance = m.moj.abun.time,
  Moj_Richness  = m.moj.rich.time,
  Moj_Diversity = m.moj.shan.time
)

# Pull out model results to create a table

# Function to extract coefficients for lmrob or glmmTMB
extract_coefs <- function(model, response_name) {
  if (inherits(model, "lmrob")) {
    # tidy() works for lmrob
    tidy(model) %>%
      mutate(Model = response_name) %>%
      select(Model, term, estimate, std.error, statistic, p.value)
  } else if (inherits(model, "glmmTMB")) {
    # For glmmTMB, use tidy with effects="fixed"
    broom.mixed::tidy(model, effects = "fixed", conf.int = FALSE) %>%
      mutate(Model = response_name) %>%
      select(Model, term, estimate, std.error, statistic, p.value)
  } else {
    stop("Unknown model type")
  }
}

# Apply the function to all models
coef_table <- imap_dfr(models, ~extract_coefs(.x, .y)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))  # round numbers for readability

# View the table
coef_table

model_table <- coef_table
write.csv(model_table, "Tables/Rodents_Fire_Time.csv")

```




## 2.2 Changes in rodent metrics over time graphs

```{r, fig.height=8, fig.width=10}


p.gb.abun.time <- ggplot(data = data_gb, aes(x= Year, y=Abun, color = Fire))+
 stat_summary(geom = "point", fun = 'mean', size = 3) +
  stat_summary(geom = "line", fun = "mean",  linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic()+
  xlab("Year")+
  ylab("Abundance")+
  coord_cartesian(ylim=c(0, 8))+
  scale_color_manual(values = c("U" = "deepskyblue3", "B"= "lightcoral"),label = c("Unburned", "Burned"))+
   scale_x_continuous(breaks = seq(min(data_gb$Year), max(data_gb$Year), by = 3)) +  # Modify x-axis breaks
   common_theme +
  theme(legend.position = "none")

p.gb.abun.avg<- ggplot(data_gb, aes(x = Fire, y = Abun, color = "black", fill = Fire)) +
  stat_summary(fun = mean, geom = "bar", width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black") +
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0,8))+
  theme_classic() +
  xlab("Mean") +
  ylab("Mean Abundance") +
  scale_color_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  scale_fill_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  common_theme+
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.text.x = element_blank()
  )


p.gb.abun.time<- p.gb.abun.time + p.gb.abun.avg +
  plot_layout(widths = c(3, 1))

p.gb.rich.time <- ggplot(data = data_gb, aes(x=Year, y= Rich, color = Fire))+
 stat_summary(geom = "point", fun = 'mean',  size = 3) +
  stat_summary(geom = "line", fun = "mean", linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic()+
  xlab("Year")+
  ylab("Species Richness")+
  coord_cartesian(ylim=c(0, 3))+
  scale_color_manual(values = c("U" = "deepskyblue3", "B"= "lightcoral"), label = c("Unburned", "Burned"))+
   scale_x_continuous(breaks = seq(min(data_gb$Year), max(data_gb$Year), by = 3)) +  # Modify x-axis breaks
   common_theme+
  theme(legend.position = "none")

p.gb.rich.avg<- ggplot(data_gb, aes(x = Fire, y = Rich, color = "black", fill = Fire)) +
  stat_summary(fun = mean, geom = "bar", width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black") +
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0,3))+
  theme_classic() +
  xlab("Mean") +
  ylab("Mean Richness") +
  scale_color_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  scale_fill_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  common_theme+
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.text.x = element_blank()
  )


p.gb.rich.time <- p.gb.rich.time + p.gb.rich.avg +
  plot_layout(widths = c(4, 1))


p.gb.shan.time <- ggplot(data = data_gb, aes(x=Year, y= Div, color = Fire))+
 stat_summary(geom = "point", fun = 'mean',  size = 3) +
  stat_summary(geom = "line", fun = "mean",  linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic()+
  xlab("Year")+
  ylab("Shannon's Diversity")+
  coord_cartesian(ylim=c(0, 1))+
  scale_color_manual(values = c("U" = "deepskyblue3", "B"= "lightcoral"),label = c("Unburned",  "Burned"))+
     scale_x_continuous(breaks = seq(min(data_gb$Year), max(data_gb$Year), by = 3)) +  # Modify x-axis breaks
   common_theme+
  theme(legend.position = "none")

p.gb.shan.avg<- ggplot(data_gb, aes(x = Fire, y = Div, color = "black", fill = Fire)) +
  stat_summary(fun = mean, geom = "bar", width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black") +
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0,1))+
  theme_classic() +
  xlab("Mean") +
  ylab("Mean Shannon's Diversity") +
  scale_color_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  scale_fill_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  common_theme+
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.text.x = element_blank()
  )


p.gb.shan.time<- p.gb.shan.time + p.gb.shan.avg +
  plot_layout(widths = c(4, 1))



p.gb.time.comb <- ggarrange(p.gb.abun.time, p.gb.rich.time, p.gb.shan.time,
                    labels = "AUTO",
                    ncol = 1, nrow = 3,
                    common.legend = TRUE,
                    legend = "right",
                    font.label=list(color="black",size=12, family = "serif"))




```


```{r, fig.height=8, fig.width=10}


p.moj.abun.time <- ggplot(data = data_moj, aes(x= Year, y=Abun, color = Fire))+
 stat_summary(geom = "point", fun = 'mean', size = 3) +
  stat_summary(geom = "line", fun = "mean",  linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic()+
  xlab("Year")+
  ylab("Abundance")+
  coord_cartesian(ylim=c(0, 8))+
  scale_color_manual(values = c("U" = "deepskyblue3", "B"= "lightcoral"),label = c("Unburned", 
                            "Burned"))+
   scale_x_continuous(breaks = seq(min(data_moj$Year), max(data_moj$Year), by = 3)) +  # Modify x-axis breaks
   common_theme+
  theme(legend.position = "none")

p.moj.abun.avg<- ggplot(data_moj, aes(x = Fire, y = Abun, color = "black", fill = Fire)) +
  stat_summary(fun = mean, geom = "bar", width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black") +
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0,8))+
  theme_classic() +
  xlab("Mean") +
  ylab("Mean Abundance") +
  scale_color_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  scale_fill_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  common_theme+
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.text.x = element_blank()
  )


p.moj.abun.time<- p.moj.abun.time + p.moj.abun.avg +
  plot_layout(widths = c(4, 1))

p.moj.rich.time <- ggplot(data = data_moj, aes(x=Year, y= Rich, color = Fire))+
 stat_summary(geom = "point", fun = 'mean',  size = 3) +
  stat_summary(geom = "line", fun = "mean", linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic()+
  xlab("Year")+
  ylab("Species Richness")+
  coord_cartesian(ylim=c(0, 3))+
  scale_color_manual(values = c("U" = "deepskyblue3", "B"= "lightcoral"), label = c("Unburned", 
                            "Burned"))+
   scale_x_continuous(breaks = seq(min(data_moj$Year), max(data_moj$Year), by = 3)) +  # Modify x-axis breaks
   common_theme+
  theme(legend.position = "none")

p.moj.rich.avg<- ggplot(data_moj, aes(x = Fire, y = Rich, color = "black", fill = Fire)) +
  stat_summary(fun = mean, geom = "bar", width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black") +
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0,3))+
  theme_classic() +
  xlab("Mean") +
  ylab("Mean Richness") +
  scale_color_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  scale_fill_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  common_theme+
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.text.x = element_blank())


p.moj.rich.time<- p.moj.rich.time + p.moj.rich.avg +
  plot_layout(widths = c(4, 1))


p.moj.shan.time <- ggplot(data = data_moj, aes(x=Year, y= Div, color = Fire))+
 stat_summary(geom = "point", fun = 'mean',  size = 3) +
  stat_summary(geom = "line", fun = "mean",  linewidth = 1.5) +
    stat_summary(geom = "errorbar", fun.data = mean_se, 
               width = 0.2) +
  theme_classic()+
  xlab("Year")+
  ylab("Shannon's Diversity")+
  coord_cartesian(ylim=c(0, 1))+
  scale_color_manual(values = c("U" = "deepskyblue3", "B"= "lightcoral"), label = c("Unburned", 
                            "Burned"))+
     scale_x_continuous(breaks = seq(min(data_moj$Year), max(data_moj$Year), by = 3)) +  # Modify x-axis breaks
   common_theme+
  theme(legend.position = "none")

p.moj.shan.avg<- ggplot(data_moj, aes(x = Fire, y = Div, color = "black", fill = Fire)) +
  stat_summary(fun = mean, geom = "bar", width = 0.5) +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2, color = "black") +
  scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim = c(0,1))+
  theme_classic() +
  xlab("Mean") +
  ylab("Mean Shannon's Diversity") +
  scale_color_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  scale_fill_manual(
    values = c("U" = "deepskyblue3", "B" = "lightcoral"),
    labels = c("U" = "Unburned", "B" = "Burned")
  ) +
  common_theme+
  theme(
    axis.title.y = element_blank(),
    axis.text.y  = element_blank(),
    axis.text.x = element_blank()
  )

p.moj.shan.time<- p.moj.shan.time + p.moj.shan.avg +
  plot_layout(widths = c(4, 1))


p.moj.time.comb <- ggarrange(p.moj.abun.time, p.moj.rich.time, p.moj.shan.time,
                    labels = "AUTO",
                    ncol = 1, nrow = 3,
                    common.legend = TRUE,
                    legend = "right",
                    font.label=list(color="black",size=12, family = "serif"))




```

Join the two plots sites to side by side

```{r, fig.width=8, fig.height=6}

# Combine Great Basin (left) and Mojave Desert (right)
p.time.all <- ggarrange(
  p.gb.abun.time, p.moj.abun.time,
  p.gb.rich.time, p.moj.rich.time,
  p.gb.shan.time, p.moj.shan.time,
  nrow = 3, ncol = 2, 
  common.legend = TRUE,
  legend = "right",
  align = "hv"
)

top_row <- ggdraw() +
  draw_label("Great Basin", x = 0.25, y = 0.5, fontfamily = "serif", size = 14) +
  draw_label("Mojave",     x = 0.675, y = 0.5, fontfamily = "serif", size = 14)

p.time.all <- plot_grid(top_row, p.time.all, ncol = 1, rel_heights = c(0.06, 1))

p.time.all

ggsave("Plots/Fire_Time_Rodents_Plot.png",p.time.all, dpi = 300, width = 8, height = 6, units = "in")


```

Results Interpretation

Abundance in the Great Basin is initially similar between burned and unburned areas, then unburned areas maintain higher abundance for around 10 years when burned plots have higher abundance in 2022

Abundance in the Mojave Desert is similar between burned and unburned plots across the study period

Species richness in the Great Basin is higher in unburned areas than burned areas for some but not all years

Species richness in the Mojave is generally slightly higher in unburned than burned plots

Shannon's diversity in the Great Basin is typically higher in unburned areas but the difference between burned and unburned areas varies between years

Shannon's diversity in the Mojave s typically higher in unburned areas but the difference between burned and unburned areas varies between years

# 3. Model Rodent responses to Fire and Climate


## 3.1 Random Forest for variable selection

```{r}

# Scale data
gbscaled <- data_gb
gbscaled[,16:35] <- scale(gbscaled[,16:35])

mojscaled <- data_moj
mojscaled[,15:34] <- scale(mojscaled[,15:34])

# Great Basin Random Forests

## Abundance

rf.gb.abun <- randomForest(
  Abun ~ Fire + TAP + MAT + SP_Spring + SP_Winter + SP_Fall + SP_Summer + MST_Spring + MST_Winter + MST_Fall + MST_Summer + lag_TAP + 
    lag_MAT + lag_SP_Spring + lag_SP_Winter + lag_SP_Fall + lag_SP_Summer + lag_MST_Spring + lag_MST_Winter + lag_MST_Fall + lag_MST_Summer,
  data = gbscaled,
  importance = TRUE,
  ntree = 500
)

## Richness

rf.gb.rich <- randomForest(
  Rich ~ Fire + TAP + MAT + SP_Spring + SP_Winter + SP_Fall + SP_Summer + MST_Spring + MST_Winter + MST_Fall + MST_Summer + lag_TAP + 
    lag_MAT + lag_SP_Spring + lag_SP_Winter + lag_SP_Fall + lag_SP_Summer + lag_MST_Spring + lag_MST_Winter + lag_MST_Fall + lag_MST_Summer,
  data = gbscaled,
  importance = TRUE,
  ntree = 500
)

## Shannon's Diversity

rf.gb.shan <- randomForest(
  Div ~ Fire + TAP + MAT + SP_Spring + SP_Winter + SP_Fall + SP_Summer + MST_Spring + MST_Winter + MST_Fall + MST_Summer + lag_TAP + 
    lag_MAT + lag_SP_Spring + lag_SP_Winter + lag_SP_Fall + lag_SP_Summer + lag_MST_Spring + lag_MST_Winter + lag_MST_Fall + lag_MST_Summer,
  data = gbscaled,
  importance = TRUE,
  ntree = 500
)

# Mojave

## Abundance

rf.moj.abun <- randomForest(
  Abun ~ Fire + TAP + MAT + SP_Spring + SP_Winter + SP_Fall + SP_Summer + MST_Spring + MST_Winter + MST_Fall + MST_Summer + lag_TAP + 
    lag_MAT + lag_SP_Spring + lag_SP_Winter + lag_SP_Fall + lag_SP_Summer + lag_MST_Spring + lag_MST_Winter + lag_MST_Fall + lag_MST_Summer,
  data = mojscaled,
  importance = TRUE,
  ntree = 500
)

## Richness

rf.moj.rich <- randomForest(
  Rich ~ Fire + TAP + MAT + SP_Spring + SP_Winter + SP_Fall + SP_Summer + MST_Spring + MST_Winter + MST_Fall + MST_Summer + lag_TAP + 
    lag_MAT + lag_SP_Spring + lag_SP_Winter + lag_SP_Fall + lag_SP_Summer + lag_MST_Spring + lag_MST_Winter + lag_MST_Fall + lag_MST_Summer,
  data = mojscaled,
  importance = TRUE,
  ntree = 500
)

##Shannon's Diversity

rf.moj.shan <- randomForest(
  Div ~ Fire + TAP + MAT + SP_Spring + SP_Winter + SP_Fall + SP_Summer + MST_Spring + MST_Winter + MST_Fall + MST_Summer + lag_TAP + 
    lag_MAT + lag_SP_Spring + lag_SP_Winter + lag_SP_Fall + lag_SP_Summer + lag_MST_Spring + lag_MST_Winter + lag_MST_Fall + lag_MST_Summer,
  data = mojscaled,
  importance = TRUE,
  ntree = 500
)


```

Plots

```{r, fig.width=8, fig.height=6}


# ---- Extract importance function ----
extract_imp <- function(rf_model, region, response) {
  randomForest::importance(rf_model, type = 1) |> 
    as.data.frame() |>
    tibble::rownames_to_column("Variable") |>
    rename(Importance = `%IncMSE`) |>
    mutate(
      Region = region,
      Response = response
    )
}

# ---- Combine importance data ----
imp_all <- bind_rows(
  extract_imp(rf.gb.abun, "Great Basin", "Abundance"),
  extract_imp(rf.gb.rich, "Great Basin", "Richness"),
  extract_imp(rf.gb.shan, "Great Basin", "Diversity"),
  extract_imp(rf.moj.abun, "Mojave", "Abundance"),
  extract_imp(rf.moj.rich, "Mojave", "Richness"),
  extract_imp(rf.moj.shan, "Mojave", "Diversity")
)

# ---- Normalize importance (optional for comparability) ----
imp_all <- imp_all %>%
  group_by(Region, Response) %>%
  mutate(RelImportance = Importance / sum(Importance)) %>%
  ungroup()

# ---- Make beehive plots ----
make_beehive <- function(region, response) {
  imp_all %>%
    filter(Region == region, Response == response) %>%
    ggplot(aes(x = RelImportance, y = reorder(Variable, RelImportance))) +
    geom_quasirandom(color = ifelse(region == "Great Basin", "#3366CC", "#CC6600"), alpha = 0.7) +
    labs(
      x = "Relative Importance",
      y = NULL,
      title = paste(region, response)
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.y = element_text(size = 9)
    )
}

# ---- Create six plots ----
p1 <- make_beehive("Great Basin", "Abundance")
p2 <- make_beehive("Mojave", "Abundance")
p3 <- make_beehive("Great Basin", "Richness")
p4 <- make_beehive("Mojave", "Richness")
p5 <- make_beehive("Great Basin", "Diversity")
p6 <- make_beehive("Mojave", "Diversity")
p1
p2
p3
p4
p5
p6

# ---- Combine in 3x2 layout ----
rf_combined <- (p1 + p2) / (p3 + p4) / (p5 + p6)

# Save as PNG
ggsave("Plots/beehive_plots.png", plot = rf_combined,
       width = 8, height = 10, dpi = 300)


```




## 3.2 Linear model selection from a priori hypotheses

# Check variable correlation first

```{r}

gbmat <- data_gb[,16:35]


cor(gbmat, method = "pearson", use = "pairwise.complete.obs")
corrplot::corrplot(cor(gbmat, method = "pearson", use = "pairwise.complete.obs"))
# Compute correlation matrix
cormat <- cor(gbmat, method = "pearson", use = "pairwise.complete.obs")

# Convert to data frame of pairs
high_cor <- as.data.frame(as.table(cormat))

# Filter for |correlation| > 0.7 and remove self-correlations
high_cor <- high_cor[abs(high_cor$Freq) > 0.7 & high_cor$Var1 != high_cor$Var2, ]

# Remove duplicate pairs (Var1-Var2 and Var2-Var1)
high_cor <- high_cor[!duplicated(t(apply(high_cor[,1:2], 1, sort))), ]

# Sort by correlation strength
high_cor <- high_cor[order(-abs(high_cor$Freq)), ]

high_cor

```

```{r}

mojmat <- data_moj[,15:34]

cor(mojmat, method = "pearson", use = "pairwise.complete.obs")
corrplot::corrplot(cor(mojmat, method = "pearson", use = "pairwise.complete.obs"))
# Compute correlation matrix
cormat <- cor(mojmat, method = "pearson", use = "pairwise.complete.obs")

# Convert to data frame of pairs
high_cor <- as.data.frame(as.table(cormat))

# Filter for |correlation| > 0.7 and remove self-correlations
high_cor <- high_cor[abs(high_cor$Freq) > 0.7 & high_cor$Var1 != high_cor$Var2, ]

# Remove duplicate pairs (Var1-Var2 and Var2-Var1)
high_cor <- high_cor[!duplicated(t(apply(high_cor[,1:2], 1, sort))), ]

# Sort by correlation strength
high_cor <- high_cor[order(-abs(high_cor$Freq)), ]

high_cor

# Don't include lag_SP_Winter and lag_MST_Spring in the same model

```


### Great Basin
### Abundance

```{r}


gbscaled%>%
  group_by(PLOT) %>%
  summarise(mean_abun = mean(Abun, na.rm = TRUE))

m.gb.abun.null <- glmmTMB(data = gbscaled, Abun ~ 1+ (1|Block),
                        family = tweedie(link = "log")) # Null Model


m.gb.abun.fire <- glmmTMB(data = gbscaled, Abun ~ Fire + (1|Block),
                        family = tweedie(link = "log")) # Hyp: Fire is the main driver of rodent abundance through vegetation removal

m.gb.abun.current <- glmmTMB(data = gbscaled, Abun ~ TAP + MAT+ (1|Block),
                        family = tweedie(link = "log"))    # Hyp: Current weather conditions drive rodent abundance

m.gb.abun.lag <- glmmTMB(data = gbscaled, Abun ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))                       # Hyp: Previous years weather drives rodent abundance

m.gb.abun.ppt <- glmmTMB(data = gbscaled, Abun ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) # Hyp: Precipitation drives rodent abundance

m.gb.abun.temp <- glmmTMB(data = gbscaled, Abun ~ MAT + MST_Spring + MST_Summer + 
                            MST_Fall + MST_Winter+ (1|Block),
                        family = tweedie(link = "log"))  # Hyp: Temperature drives rodent abundance

m.gb.abun.fire.ppt <- glmmTMB(data = gbscaled, Abun ~ TAP * Fire+ (1|Block),
                        family = tweedie(link = "log"))       # Hyp: Interaction between fire and precipitation drives rodent abundance

m.gb.abun.fire.temp <- glmmTMB(data = gbscaled, Abun ~ MAT * Fire+ (1|Block),
                        family = tweedie(link = "log"))           # Hyp: The interaction between fire and temperature drives rodent abundance

m.gb.abun.rf <- glmmTMB(data = gbscaled,
                        Abun ~ SP_Spring * Fire + (1|Block),
                        family = tweedie(link = "log")) # Highest variable importance from rf with fire  


m.gb.abun.sel <- model.sel(m.gb.abun.null, m.gb.abun.fire, m.gb.abun.current, m.gb.abun.lag, m.gb.abun.rf, 
                           m.gb.abun.ppt, m.gb.abun.temp, m.gb.abun.fire.ppt, m.gb.abun.fire.temp)

m.gb.abun.sel
summary(m.gb.abun.rf)

# Extract top model
gb.abun.top_model <- get.models(m.gb.abun.sel, 1)[[1]]

# Get top model coefficients
gb.abun.coef_table <- broom.mixed::tidy(gb.abun.top_model, effects = "fixed", conf.int = TRUE) |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))

# View it
gb.abun.coef_table

p.gb.abun <- ggplot(gb.abun.coef_table, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 12) +
  labs(
    x = "Coefficient Estimate (+/- 95% CI)",
    y = "Abundance",
  ) + theme_classic()+ common_theme

p.gb.abun

```

Model check

```{r}

sim_res <- simulateResiduals(m.gb.abun.rf, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

```

### Richness

```{r}


m.gb.rich.null <- glmmTMB(data = gbscaled, Rich ~ 1+ (1|Block),
                        family = tweedie(link = "log")) # Null Model


m.gb.rich.fire <- glmmTMB(data = gbscaled, Rich ~ Fire + (1|Block),
                        family = tweedie(link = "log")) # Hyp: Fire is the main driver of rodent richness through vegetation removal

m.gb.rich.current <- glmmTMB(data = gbscaled, Rich ~ TAP + MAT+ (1|Block),
                        family = tweedie(link = "log"))    # Hyp: Current weather conditions drive rodent richness

m.gb.rich.lag <- glmmTMB(data = gbscaled, Rich ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))                       # Hyp: Previous years weather drives rodent richness

m.gb.rich.ppt <- glmmTMB(data = gbscaled, Rich ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) # Hyp: Precipitation drives rodent richness

m.gb.rich.temp <- glmmTMB(data = gbscaled, Rich ~ MAT + MST_Spring + MST_Summer + 
                            MST_Fall + MST_Winter+ (1|Block),
                        family = tweedie(link = "log"))  # Hyp: Temperature drives rodent richness

m.gb.rich.fire.ppt <- glmmTMB(data = gbscaled, Rich ~ TAP * Fire+ (1|Block),
                        family = tweedie(link = "log"))             # Hyp: Interaction between fire and precipitation drives rodent richness

m.gb.rich.fire.temp <- glmmTMB(data = gbscaled, Rich ~ MAT * Fire+ (1|Block),
                        family = tweedie(link = "log"))           # Hyp: The interaction between fire and temperature drives rodent richness

m.gb.rich.rf <- glmmTMB(data = gbscaled, Rich ~ lag_SP_Spring * Fire+ (1|Block),
                        family = tweedie(link = "log")) # Highest variable importance from rf with fire  


m.gb.rich.sel <- model.sel(m.gb.rich.null, m.gb.rich.fire, m.gb.rich.current, m.gb.rich.lag, m.gb.rich.rf, 
                           m.gb.rich.ppt, m.gb.rich.temp, m.gb.rich.fire.ppt, m.gb.rich.fire.temp)

m.gb.rich.sel

gb.rich.top_model <- get.models(m.gb.rich.sel, 1)[[1]]
gb.rich.coef_table <-broom.mixed::tidy(gb.rich.top_model, effects = "fixed", conf.int = TRUE) |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))

# View it
gb.rich.coef_table

p.gb.rich <- ggplot(gb.rich.coef_table, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 12) +
  labs(
    x = "Coefficient Estimate (+/- 95% CI)",
    y = "Species Richness",
  ) + theme_classic()+ common_theme

p.gb.rich

```

Model check

```{r}

sim_res <- simulateResiduals(m.gb.rich.rf, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

```
### Shannon's Diversity

```{r}


m.gb.shan.null <- glmmTMB(data = gbscaled, Div ~ 1+ (1|Block),
                        family = tweedie(link = "log")) # Null Model


m.gb.shan.fire <- glmmTMB(data = gbscaled, Div ~ Fire + (1|Block),
                        family = tweedie(link = "log")) # Hyp: Fire is the main driver of rodent Shannon's Diversity through vegetation removal

m.gb.shan.current <- glmmTMB(data = gbscaled, Div ~ TAP + MAT+ (1|Block),
                        family = tweedie(link = "log"))    # Hyp: Current weather conditions drive rodent Shannon's Diversity

m.gb.shan.lag <- glmmTMB(data = gbscaled, Div ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))                       # Hyp: Previous years weather drives rodent Shannon's Diversity

m.gb.shan.ppt <- glmmTMB(data = gbscaled, Div ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) # Hyp: Precipitation drives rodent Shannon's Diversity

m.gb.shan.temp <- glmmTMB(data = gbscaled, Div ~ MAT + MST_Spring + MST_Summer + 
                            MST_Fall + MST_Winter+ (1|Block),
                        family = tweedie(link = "log"))  # Hyp: Temperature drives rodent Shannon's Diversity

m.gb.shan.fire.ppt <- glmmTMB(data = gbscaled, Div ~ TAP * Fire+ (1|Block),
                        family = tweedie(link = "log"))             # Hyp: Interaction between fire and precipitation drives rodent Shannon's Diversity

m.gb.shan.fire.temp <- glmmTMB(data = gbscaled, Div ~ MAT * Fire+ (1|Block),
                        family = tweedie(link = "log"))           # Hyp: The interaction between fire and temperature drives rodent Shannon's Diversity

m.gb.shan.rf <- glmmTMB(data = gbscaled, Div ~ SP_Spring * Fire+ (1|Block),
                        family = tweedie(link = "log")) # Highest variable importance from rf with fire  


m.gb.shan.sel <- model.sel(m.gb.shan.null, m.gb.shan.fire, m.gb.shan.current, m.gb.shan.lag, m.gb.shan.rf, 
                           m.gb.shan.ppt, m.gb.shan.temp, m.gb.shan.fire.ppt, m.gb.shan.fire.temp)

m.gb.shan.sel

gb.shan.top_model <- get.models(m.gb.shan.sel, 1)[[1]]
gb.shan.coef_table <- broom.mixed::tidy(gb.shan.top_model, effects = "fixed", conf.int = TRUE) |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))

# View it
gb.shan.coef_table

p.gb.shan <- ggplot(gb.shan.coef_table, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 12) +
  labs(
    x = "Coefficient Estimate (+/- 95% CI)",
    y = "Shannon's Diversity",
  ) + theme_classic()+ common_theme

p.gb.shan

```

Model check

```{r}

sim_res <- simulateResiduals(m.gb.shan.lag, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

```

### Mojave
### Abundance

```{r}


m.moj.abun.null <- glmmTMB(data = mojscaled, Abun ~ 1+ (1|Block),
                        family = tweedie(link = "log")) # Null Model


m.moj.abun.fire <- glmmTMB(data = mojscaled, Abun ~ Fire + (1|Block),
                        family = tweedie(link = "log")) # Hyp: Fire is the main driver of rodent abundance through vegetation removal

m.moj.abun.current <- glmmTMB(data = mojscaled, Abun ~ TAP + MAT+ (1|Block),
                        family = tweedie(link = "log"))    # Hyp: Current weather conditions drive rodent abundance

m.moj.abun.lag <- glmmTMB(data = mojscaled, Abun ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))                       # Hyp: Previous years weather drives rodent abundance

# m.moj.abun.ppt <- glmmTMB(data = mojscaled, Abun ~ TAP + SP_Spring + SP_Summer+
#                            SP_Fall + SP_Winter+ (1|Block),
#                         family = tweedie(link = "log")) # Hyp: Precipitation drives rodent abundance


# We removed uninformative parameters (see bottom) and this is the updated best model

m.moj.abun.ppt <- glmmTMB(data = mojscaled, Abun ~ TAP +  SP_Summer+
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

m.moj.abun.temp <- glmmTMB(data = mojscaled, Abun ~ MAT + MST_Spring + MST_Summer + 
                            MST_Fall + MST_Winter+ (1|Block),
                        family = tweedie(link = "log"))  # Hyp: Temperature drives rodent abundance

m.moj.abun.fire.ppt <- glmmTMB(data = mojscaled, Abun ~ TAP * Fire+ (1|Block),
                        family = tweedie(link = "log"))             # Hyp: Interaction between fire and precipitation drives rodent abundance

m.moj.abun.fire.temp <- glmmTMB(data = mojscaled, Abun ~ MAT * Fire+ (1|Block),
                        family = tweedie(link = "log"))           # Hyp: The interaction between fire and temperature drives rodent abundance

m.moj.abun.rf <- glmmTMB(data = mojscaled, Abun ~ lag_MAT * Fire+ (1|Block),
                        family = tweedie(link = "log")) # Highest variable importance from rf with fire  


m.moj.abun.sel <- model.sel(m.moj.abun.null, m.moj.abun.fire, m.moj.abun.current, m.moj.abun.lag, m.moj.abun.rf, 
                           m.moj.abun.ppt, m.moj.abun.temp, m.moj.abun.fire.ppt, m.moj.abun.fire.temp)

m.moj.abun.sel

moj.abun.top_model <- get.models(m.moj.abun.sel, 1)[[1]]
moj.abun.coef_table <- broom.mixed::tidy(moj.abun.top_model, effects = "fixed", conf.int = TRUE) |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))

# View it
moj.abun.coef_table

p.moj.abun <- ggplot(moj.abun.coef_table, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 12) +
  labs(
    x = "Coefficient Estimate (+/- 95% CI)",
    y = NULL,
  ) + theme_classic()+ common_theme

p.moj.abun

```

Model check

```{r}

sim_res <- simulateResiduals(m.moj.abun.ppt, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

# ── Step 1: Estimate c-hat from your most complex model ──────────────────────
# Use your most parameterized model as the "global" model
# m.moj.shan.temp has the most parameters among your single-theme models

sim_global <- simulateResiduals(m.moj.abun.ppt, n = 1000)
disp_test  <- testDispersion(sim_global, plot = TRUE)

c_hat <- disp_test$statistic  # ratio of observed/expected variance = c-hat
cat("c-hat:", round(c_hat, 3), "\n")

# ── Step 2: QAIC function for glmmTMB ────────────────────────────────────────
# Formula: QAIC = (-2 * logLik) / c_hat + 2 * K
# attr(logLik, "df") gives K (all estimated params incl. random effects + dispersion)

calc_QAIC <- function(model, c_hat) {
  ll <- as.numeric(logLik(model))
  k  <- attr(logLik(model), "df")
  (-2 * ll) / c_hat + 2 * k
}

# ── Step 3: Build comparison table ───────────────────────────────────────────
models <- list(
  null       = m.moj.abun.null,
  fire       = m.moj.abun.fire,
  current    = m.moj.abun.current,
  lag        = m.moj.abun.lag,
  ppt        = m.moj.abun.ppt,
  temp       = m.moj.abun.temp,
  fire.ppt   = m.moj.abun.fire.ppt,
  fire.temp  = m.moj.abun.fire.temp,
  rf         = m.moj.abun.rf
)

qaic_table <- data.frame(
  Model    = names(models),
  K        = sapply(models, \(m) attr(logLik(m), "df")),
  LogLik   = sapply(models, \(m) round(as.numeric(logLik(m)), 2)),
  QAIC     = sapply(models, \(m) round(calc_QAIC(m, c_hat), 2))
)

qaic_table$delta_QAIC <- round(qaic_table$QAIC - min(qaic_table$QAIC), 2)

# Akaike weights from QAIC
qaic_table$weight <- round(
  exp(-0.5 * qaic_table$delta_QAIC) / sum(exp(-0.5 * qaic_table$delta_QAIC)), 3
)

qaic_table <- qaic_table[order(qaic_table$QAIC), ]
print(qaic_table)
write.csv(qaic_table, "Tables/qAIC_Mojave_Abundance.csv", row.names = FALSE)

```

### Richness

```{r}


m.moj.rich.null <- glmmTMB(data = mojscaled, Rich ~ 1+ (1|Block),
                        family = tweedie(link = "log")) # Null Model


m.moj.rich.fire <- glmmTMB(data = mojscaled, Rich ~ Fire+ (1|Block),
                        family = tweedie(link = "log") ) # Hyp: Fire is the main driver of rodent richness through vegetation removal

m.moj.rich.current <- glmmTMB(data = mojscaled, Rich ~ TAP + MAT+ (1|Block),
                        family = tweedie(link = "log"))    # Hyp: Current weather conditions drive rodent richness

m.moj.rich.lag <- glmmTMB(data = mojscaled, Rich ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))                       # Hyp: Previous years weather drives rodent richness

# m.moj.rich.ppt <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + SP_Summer+
#                            SP_Fall + SP_Winter+ (1|Block),
#                         family = tweedie(link = "log")) # Hyp: Precipitation drives rodent richness

# We removed uninformative parameters SP_Fall and SP_Summer (see Bottom)

m.moj.rich.ppt <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + SP_Winter+ (1|Block),
                        family = tweedie(link = "log"))

m.moj.rich.temp <- glmmTMB(data = mojscaled, Rich ~ MAT + MST_Spring + MST_Summer + 
                            MST_Fall + MST_Winter+ (1|Block),
                        family = tweedie(link = "log"))  # Hyp: Temperature drives rodent richness

m.moj.rich.fire.ppt <- glmmTMB(data = mojscaled, Rich ~ TAP * Fire+ (1|Block),
                        family = tweedie(link = "log"))             # Hyp: Interaction between fire and precipitation drives rodent richness

m.moj.rich.fire.temp <- glmmTMB(data = mojscaled, Rich ~ MAT * Fire+ (1|Block),
                        family = tweedie(link = "log"))           # Hyp: The interaction between fire and temperature drives rodent richness

m.moj.rich.rf <- glmmTMB(data = mojscaled, Rich ~ lag_SP_Winter * Fire+ (1|Block),
                        family = tweedie(link = "log")) # Highest variable importance from rf with fire  


m.moj.rich.sel <- model.sel(m.moj.rich.null, m.moj.rich.fire, m.moj.rich.current, m.moj.rich.lag, m.moj.rich.rf, 
                           m.moj.rich.ppt, m.moj.rich.temp, m.moj.rich.fire.ppt, m.moj.rich.fire.temp)

m.moj.rich.sel

moj.rich.top_model <- get.models(m.moj.rich.sel, 1)[[1]]
moj.rich.coef_table <- broom.mixed::tidy(moj.rich.top_model, effects = "fixed", conf.int = TRUE) |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))

# View it
moj.rich.coef_table

p.moj.rich <- ggplot(moj.rich.coef_table, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 12) +
  labs(
    x = "Coefficient Estimate (+/- 95% CI)",
    y = NULL,
  ) + theme_classic()+ common_theme

p.moj.rich

```

Model check

```{r}

sim_res <- simulateResiduals(m.moj.rich.ppt, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

```
### Shannon's Diversity

```{r}


m.moj.shan.null <- glmmTMB(data = mojscaled, Div ~ 1+ (1|Block),
                        family = tweedie(link = "log")) # Null Model


m.moj.shan.fire <- glmmTMB(data = mojscaled, Div ~ Fire + (1|Block),
                        family = tweedie(link = "log")) # Hyp: Fire is the main driver of rodent Shannon's Diversity through vegetation removal

m.moj.shan.current <- glmmTMB(data = mojscaled, Div ~ TAP + MAT+ (1|Block),
                        family = tweedie(link = "log"))    # Hyp: Current weather conditions drive rodent Shannon's Diversity

m.moj.shan.lag <- glmmTMB(data = mojscaled, Div ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))                       # Hyp: Previous years weather drives rodent Shannon's Diversity

# m.moj.shan.ppt <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + SP_Summer+
#                            SP_Fall + SP_Winter+ (1|Block),
#                         family = tweedie(link = "log")) # Hyp: Precipitation drives rodent Shannon's Diversity

# We removed uninformative parameters SP_Fall and SP_Summer (see bottom)

m.moj.shan.ppt <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + 
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log"))

m.moj.shan.temp <- glmmTMB(data = mojscaled, Div ~ MAT + MST_Spring + MST_Summer + 
                            MST_Fall + MST_Winter+ (1|Block),
                        family = tweedie(link = "log"))  # Hyp: Temperature drives rodent Shannon's Diversity

m.moj.shan.fire.ppt <- glmmTMB(data = mojscaled, Div ~ TAP * Fire+ (1|Block),
                        family = tweedie(link = "log"))             # Hyp: Interaction between fire and precipitation drives rodent Shannon's Diversity

m.moj.shan.fire.temp <- glmmTMB(data = mojscaled, Div ~ MAT * Fire+ (1|Block),
                        family = tweedie(link = "log"))           # Hyp: The interaction between fire and temperature drives rodent Shannon's Diversity

m.moj.shan.rf <- glmmTMB(data = mojscaled, Div ~ lag_SP_Summer * Fire+ (1|Block),
                        family = tweedie(link = "log")) # Highest variable importance from rf with fire  


m.moj.shan.sel <- model.sel(m.moj.shan.null, m.moj.shan.fire, m.moj.shan.current, m.moj.shan.lag, m.moj.shan.rf, 
                           m.moj.shan.ppt, m.moj.shan.temp, m.moj.shan.fire.ppt, m.moj.shan.fire.temp)

m.moj.shan.sel


moj.shan.top_model <- get.models(m.moj.shan.sel, 1)[[1]]
moj.shan.coef_table <- broom.mixed::tidy(moj.shan.top_model, effects = "fixed", conf.int = TRUE) |>
  dplyr::mutate(across(where(is.numeric), ~ round(.x, 3)))

# View it
moj.shan.coef_table

p.moj.shan <- ggplot(moj.shan.coef_table, aes(x = estimate, y = reorder(term, estimate))) +
  geom_point(size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  theme_minimal(base_size = 12) +
  labs(
    x = "Coefficient Estimate (+/- 95% CI)",
    y = NULL,
  ) + theme_classic()+ common_theme

p.moj.shan

```

Model check

```{r}

sim_res <- simulateResiduals(m.moj.shan.ppt, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)


```

Underdispersed therefore do a sensitivity check

```{r}

sens.zit <- glmmTMB(Div ~ TAP + SP_Spring + SP_Winter + (1|Block),
                    data = mojscaled,
                    family = tweedie(link = "log"),
                    ziformula = ~1)

summary(sens.zit)
summary(m.moj.shan.ppt)

AIC(m.moj.shan.ppt, sens.zit)

sim_res <- simulateResiduals(sens.zit, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

```

Use a zero-inflated model instead

```{r}

m.moj.shan.null      <- glmmTMB(Div ~ 1 + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.fire      <- glmmTMB(Div ~ Fire + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.current   <- glmmTMB(Div ~ TAP + MAT + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.lag       <- glmmTMB(Div ~ lag_TAP + lag_MAT + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.ppt       <- glmmTMB(Div ~ TAP + SP_Spring + SP_Winter + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.temp      <- glmmTMB(Div ~ MAT + MST_Spring + MST_Summer + 
                                  MST_Fall + MST_Winter + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.fire.ppt  <- glmmTMB(Div ~ TAP * Fire + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.fire.temp <- glmmTMB(Div ~ MAT * Fire + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)
m.moj.shan.rf        <- glmmTMB(Div ~ lag_SP_Summer * Fire + (1|Block), 
                                 data = mojscaled, family = tweedie(link="log"), 
                                 ziformula = ~1)

# New model selection
m.moj.shan.sel <- model.sel(m.moj.shan.null, m.moj.shan.fire, m.moj.shan.current,
                             m.moj.shan.lag, m.moj.shan.rf, m.moj.shan.ppt,
                             m.moj.shan.temp, m.moj.shan.fire.ppt, 
                             m.moj.shan.fire.temp)
m.moj.shan.sel



```

Now check dispersion

```{r}

sim_res <- simulateResiduals(m.moj.shan.ppt, n = 1000)
plot(sim_res)           # QQ plot + residuals vs. fitted
testDispersion(sim_res) # Over/underdispersion test
testZeroInflation(sim_res)

```



```{r}

# Combine all tables into one
all_tables <- bind_rows(
  as.data.frame(m.gb.abun.sel),
  as.data.frame(m.gb.rich.sel),
  as.data.frame(m.gb.shan.sel),
  as.data.frame(m.moj.abun.sel),
  as.data.frame(m.moj.rich.sel),
  as.data.frame(m.moj.shan.sel)
)


write.csv(all_tables, "Tables/rawselection.csv")

```

```{r}

#Combine all top models

all_tops <- bind_rows(
  as.data.frame(gb.abun.coef_table),
  as.data.frame(gb.rich.coef_table),
  as.data.frame(gb.shan.coef_table),
  as.data.frame(moj.abun.coef_table),
  as.data.frame(moj.rich.coef_table),
  as.data.frame(moj.shan.coef_table)
)

write.csv(all_tops, "Tables/rawtopmodels.csv")

```




Check for uninformative parameters

### Great Basin Abundance

No Uninformative parameters as the top model is a single interaction term

### Great Basin Richness

No Uninformative parameters as the top model is a single interaction term

### Great Basin Shannon's Diversity

```{r}

m.gb.shan.lag <- glmmTMB(data = gbscaled, Div ~ lag_TAP + lag_MAT+ (1|Block),
                        family = tweedie(link = "log"))         

a1 <- glmmTMB(data = gbscaled, Div ~ lag_TAP +  (1|Block),
                        family = tweedie(link = "log"))  

a2 <- glmmTMB(data = gbscaled, Div ~ lag_MAT +  (1|Block),
                        family = tweedie(link = "log"))   

model.sel(m.gb.shan.lag, a1, a2)

```



### Mojave Abundance

```{r}

m.moj.abun.ppt <- glmmTMB(data = mojscaled, Abun ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a1 <- glmmTMB(data = mojscaled, Abun ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall +  (1|Block),
                        family = tweedie(link = "log")) 

a2 <- glmmTMB(data = mojscaled, Abun ~ TAP + SP_Spring + SP_Summer+
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a3 <- glmmTMB(data = mojscaled, Abun ~ TAP + SP_Spring + 
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a4 <- glmmTMB(data = mojscaled, Abun ~ TAP + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a5 <- glmmTMB(data = mojscaled, Abun ~ SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a6 <- glmmTMB(data = mojscaled, Abun ~ TAP +  SP_Summer+
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

# Use this to test all the remaining removed explanatory variables
a7 <- glmmTMB(data = mojscaled, Abun ~ SP_Summer +  SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) # Concludes that a6 had the best model 


model.sel(m.moj.abun.ppt,a1,a2,a3,a4,a5, a6, a7)

# We go back to the Mojave Abundance model selection and update it to include the updated model from a6 as sp_spring and sp_fall are
# uninformative parameters


```
### Mojave Richness

```{r}

m.moj.rich.ppt <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a1 <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall +  (1|Block),
                        family = tweedie(link = "log")) 

a2 <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + SP_Summer+
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a3 <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + 
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a4 <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a5 <- glmmTMB(data = mojscaled, Rich ~ SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a6 <- glmmTMB(data = mojscaled, Rich ~ TAP + SP_Spring + 
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

# Use this to test all the remaining removed explanatory variables
a7 <- glmmTMB(data = mojscaled, Rich ~ SP_Spring + SP_Winter +  (1|Block),
                        family = tweedie(link = "log")) # a6 is best model


model.sel(m.moj.rich.ppt,a1,a2,a3,a4,a5, a6, a7)

# Update top richness  model above to include the model from a6 removing the uninformative parameters SP_Fall and SP_Summer


```
### Mojave Shannon's Diversity

```{r}

m.moj.shan.ppt <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a1 <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + SP_Summer+
                           SP_Fall +  (1|Block),
                        family = tweedie(link = "log")) 

a2 <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + SP_Summer+
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a3 <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + 
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a4 <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a5 <- glmmTMB(data = mojscaled, Div ~ SP_Spring + SP_Summer+
                           SP_Fall + SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

a6 <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring + 
                           SP_Winter+ (1|Block),
                        family = tweedie(link = "log")) 

# Use this to test all the remaining removed explanatory variables
a7 <- glmmTMB(data = mojscaled, Div ~ TAP + SP_Spring +  (1|Block),
                        family = tweedie(link = "log")) # a6 is best model


model.sel(m.moj.shan.ppt,a1,a2,a3,a4,a5, a6, a7)

# Update top diversity model above to include the model from a6 removing the uninformative parameters SP_Fall and SP_Summer


```

Compile model results

```{r}

# Remove x-axis titles from all except the bottom plots
p.gb.abun <- p.gb.abun + theme(axis.title.x = element_blank())
p.moj.abun <- p.moj.abun + theme(axis.title.x = element_blank())
p.gb.rich <- p.gb.rich + theme(axis.title.x = element_blank())
p.moj.rich <- p.moj.rich + theme(axis.title.x = element_blank())

combined <- ggarrange(
  p.gb.abun, p.moj.abun,
  p.gb.rich, p.moj.rich,
  p.gb.shan, p.moj.shan,
  nrow = 3, ncol = 2,
  common.legend = TRUE,
  legend = "right",
  align = "hv"
)



top_row <- ggdraw() +
  draw_label("Great Basin", x = 0.35, y = 0.5, fontfamily = "serif", size = 16) +
  draw_label("Mojave",     x = 0.8, y = 0.5, fontfamily = "serif", size = 16)

final_plot <- plot_grid(top_row, combined, ncol = 1, rel_heights = c(0.06, 1))


final_plot




ggsave("Plots/Fire_Climate_Rodents_Plot.png",final_plot, dpi = 300, width = 10, height = 8, units = "in")


```


# 4 Species differences

## Great Basin

```{r}

head(data_gb)

species_cols <- colnames(data_gb[, 5:14])

# Function to run test + summary for one species
summ_fun <- function(sp) {

  df_sub <- data_gb %>% select(Fire, all_of(sp))

  test <- t.test(df_sub[[sp]] ~ df_sub$Fire)

  Mean_Unburned <- mean(df_sub[[sp]][df_sub$Fire == "U"], na.rm = TRUE)
  Mean_Burned   <- mean(df_sub[[sp]][df_sub$Fire == "B"], na.rm = TRUE)
  Percent_Diff  <- 100 * (Mean_Burned - Mean_Unburned) / Mean_Unburned
  N_total       <- sum(df_sub[[sp]], na.rm = TRUE)

  tibble(
    Species        = sp,
    N              = N_total,
    Mean_Unburned  = round(Mean_Unburned, 2),
    Mean_Burned    = round(Mean_Burned, 2),
    Percent_Diff   = round(Percent_Diff, 2),
    Test_Statistic = round(test$statistic, 2),
    P_Value        = round(test$p.value, 3)
  )
}

# Apply to all species and combine
gb_species <- map_df(species_cols, summ_fun)%>%
  arrange(desc(N))

gb_species

write.csv(gb_species, "Tables/GB_Rodent_Species_Changes.csv")

```

## Mojave

```{r}

head(data_moj)

species_cols <- colnames(data_moj[, 5:13])

# Function to run test + summary for one species
summ_fun <- function(sp) {

  df_sub <- data_moj %>% select(Fire, all_of(sp))

  test <- t.test(df_sub[[sp]] ~ df_sub$Fire)

  Mean_Unburned <- mean(df_sub[[sp]][df_sub$Fire == "U"], na.rm = TRUE)
  Mean_Burned   <- mean(df_sub[[sp]][df_sub$Fire == "B"], na.rm = TRUE)
  Percent_Diff  <- 100 * (Mean_Burned - Mean_Unburned) / Mean_Unburned
  N_total       <- sum(df_sub[[sp]], na.rm = TRUE)

  tibble(
    Species        = sp,
    N              = N_total,
    Mean_Unburned  = round(Mean_Unburned, 2),
    Mean_Burned    = round(Mean_Burned, 2),
    Percent_Diff   = round(Percent_Diff, 2),
    Test_Statistic = round(test$statistic, 2),
    P_Value        = round(test$p.value, 3)
  )
}

# Apply to all species and combine
moj_species <- map_df(species_cols, summ_fun)%>%
  arrange(desc(N))

moj_species <- moj_species[1:11,]
moj_species

write.csv(moj_species, "Tables/Mojave_Rodent_Species_Changes.csv")

```


# 5 Structural Equation Modeling: Invasive Species Effects

## 5.1 Add plant data

```{r}

gbplants <- read.csv("Data/Plants/GreatBasin_Annuals.csv")
mojplants <- read.csv("Data/Plants/Mojave_Data.csv")

# Select columns with identifiers and data we need (These are densities per m squared)
gbplants <- gbplants |> select(Year,Block,Plot,BRTE,ALAL, CETE,HAGL)
mojplants <- mojplants |> select(Year, Block, Plot, BRRU, ASNU, ERCI, SCAR)

# Match plot to PLOT
colnames(gbplants)[3]<- "PLOT"
colnames(mojplants)[3] <- "PLOT"

# Only keep BS and US
gbplants <- gbplants |> filter(PLOT %in% c("BS", "US"))
mojplants <- mojplants |> filter(PLOT %in% c("BS", "US"))



#scale densities
gbplants[, 4:7] <- scale(gbplants[,4:7])
mojplants[,4:7] <- scale(mojplants[,4:7])


gbplants$PLOT <- as.factor(gbplants$PLOT)
mojplants$PLOT <- as.factor(mojplants$PLOT)



  
# the rodent data needs to become annual means to match the plant data


gbsemdata <- dplyr::left_join(gbscaled, gbplants, by = c("Year", "Block", "PLOT")) |>
  select(Year, Block, PLOT, Fire, Abun, Rich, Div, TAP, MAT, BRTE, CETE, ALAL, HAGL) |>
  na.omit()

mojsemdata <- left_join(mojscaled, mojplants, by = c("Year", "Block", "PLOT")) |>
  select(Year, Block, PLOT, Fire, Abun, Rich, Div, TAP, MAT, BRRU, SCAR, ERCI, ASNU) |>
  na.omit()

```



## 5.2 SEM models

# Starting without plant data for now! replace this when I get it

### Great Basin

```{r}


gb.sem <- "
        # Rodent Response to veg
        Abun ~ BRTE + HAGL + CETE + ALAL + Fire + TAP + MAT
        Rich ~ BRTE + HAGL + CETE + ALAL + Fire + TAP + MAT
        Div ~ BRTE + HAGL + CETE + ALAL + Fire + TAP + MAT
        
        # Veg Response to Climate
        BRTE ~ Fire + TAP + MAT
        HAGL ~ Fire + TAP + MAT
        CETE ~ Fire + TAP + MAT
        ALAL ~ Fire + TAP + MAT
"


fitmas <- sem(gb.sem, data = gbsemdata)

summary(fitmas, standardized = TRUE, fit.measures = TRUE)


gb_sem_est <- standardizedSolution(fitmas)

write.csv(gb_sem_est, "Tables/GB_SEM_Results.csv", row.names = FALSE)

gb_sem_est |> filter(pvalue <=0.1)

```


### Mojave

```{r}


moj.sem <- "
        # Rodent Response to veg
        Abun ~ BRRU + SCAR + ASNU + ERCI + Fire + TAP + MAT
        Rich ~ BRRU + SCAR + ASNU + ERCI + Fire + TAP + MAT
        Div ~ BRRU + SCAR + ASNU + ERCI + Fire + TAP + MAT

        # Veg Response to Climate
        BRRU ~ Fire + TAP + MAT
        SCAR ~ Fire + TAP + MAT
        ERCI ~ Fire + TAP + MAT
        ASNU ~ Fire + TAP + MAT
"



fitmas <- sem(moj.sem, data = mojsemdata)
summary(fitmas, standardized = TRUE, fit.measures = TRUE)

moj_sem_est <- standardizedSolution(fitmas)

write.csv(moj_sem_est, "Tables/MOJ_SEM_Results.csv", row.names = FALSE)

moj_sem_est |> filter(pvalue <=0.1)

```



## 5.3 SEM plots

```{r}
library(DiagrammeR)

gbsem <- grViz("
digraph sem_new {
  graph [layout=dot, rankdir=LR, splines=line]
  node [shape=rectangle, style=filled, fillcolor=lightgrey, fontsize=10]

  # ---- Left ----
  {rank=same;
    TAP  [label='Total Annual\nPrecipitation'];
    Fire [label='Fire'];
  }

  # ---- Middle ----
  {rank=same;
    ALAL [label=<<I>A. allysoides</I><BR/>density>];
    CETE [label=<<I>C. testiculata</I><BR/>density>];
    HAGL [label=<<I>H. glomeratus</I><BR/>density>];
  }

  # ---- Right ----
  {rank=same;
    Abun [label='Abundance'];
    Rich [label='Richness'];
  }

  # ---- Invisible layout anchors ----
  TAP  -> ALAL [style=invis];
  ALAL -> Abun [style=invis];

  # Climate → Vegetation
  Fire -> CETE [color='#006400', penwidth=1.37];
  TAP  -> CETE [color='#006400', penwidth=4.00];

  Fire -> ALAL [color='#006400', penwidth=1.10];
  TAP  -> ALAL [color='#006400', penwidth=1.94];
  
  Fire -> HAGL [color='#006400', penwidth=2.00];

  # Vegetation → Abundance
  CETE -> Abun [color='#8B0000', penwidth=1.21];
  ALAL -> Abun [color='#006400', penwidth=1.16];

  # Vegetation → Richness
  HAGL -> Rich [color='#8B0000', penwidth=1.16];
  ALAL -> Rich [color='#8B0000', penwidth=1.00];

}
")

gbsem

#Mojave
library(DiagrammeR)

mojsem <- grViz("
digraph sem_sig_scaled_straight {
  graph [layout = dot, rankdir = LR, splines = line]
  node [shape = rectangle, style = filled, fillcolor = lightgrey, fontsize=12]

  # ---- Left: Climate ----
  {rank = same;
    Fire [label='Fire'];
    MAT  [label='Mean Annual\nTemperature'];
  }

  # ---- Middle: Vegetation ----
  BRRU [label=<<I>B. rubens</I><BR/>density>];

  # ---- Right: Responses ----
  Rich [label='Species\nRichness'];
  Div  [label='Diversity'];

  # ---- Invisible anchors for layout ----
  Fire -> BRRU [style=invis];
  BRRU -> Rich [style=invis];

  # ---- Climate → Vegetation ----
  Fire -> BRRU [color='#8B0000', penwidth=1.00];
  MAT  -> BRRU [color='#8B0000', penwidth=2.14];

  # ---- Vegetation → Responses ----
  BRRU -> Rich [color='#006400', penwidth=1.71];
  BRRU -> Div  [color='#006400', penwidth=1.81];
}
")
mojsem


```



```{r}

library(DiagrammeRsvg)
library(rsvg)

# Convert to SVG
gb_svg <- export_svg(gbsem)
moj_svg <- export_svg(mojsem)

# Write PNG
rsvg_png(charToRaw(gb_svg), "plots/gbsem.png", width = 2000, height = 1200)
rsvg_png(charToRaw(moj_svg), "Plots/mojsem.png", width = 2000, height = 1200)

gbsem
mojsem

```

```{r}

library(magick)
library(cowplot)

# Read PNGs
gb_img <- image_read("plots/gbsem.png")
moj_img <- image_read("Plots/mojsem.png")

# Convert magick images to ggplot objects
gb_gg <- ggdraw() + draw_image(gb_img)
moj_gg <- ggdraw() + draw_image(moj_img)

# Combine with labels closer
combined <- plot_grid(
  gb_gg, moj_gg,
  labels = c("A", "B"),
  ncol = 1,
  label_size = 20,
  label_x = 0.02,   # move labels closer to left edge
  label_y = 0.98    # move labels closer to top edge
)
combined
# Save combined figure
ggsave("plots/combined_sem.png", combined, width = 16, height = 16, dpi = 300)

```

Reburn Supplementary test

```{r}

reburn_gb <- data_gb |> filter(Year >=2016)|>
  filter(PLOT=="BS")

reburn_gb <- reburn_gb %>%
  mutate(design = case_when(
    Block %in% c(1, 4) ~ "adjacent",
    Block %in% c(2, 3, 5) ~ "across",
    TRUE ~ NA_character_
  ))

reburn_gb <- reburn_gb |> select(Year, Block, PLOT, Abun, Rich, Div, design, Time)



m.gb.abun.reb <- lmrob(data = reburn_gb, Abun ~ design)
m.gb.rich.reb <- lmrob(data = reburn_gb, Rich ~ design)
m.gb.shan.reb <- glmmTMB(data = reburn_gb, Div ~ design,
                          family = tweedie(link = "log"))

reb.models <- list(
  GB_Abundance = m.gb.abun.reb,
  GB_Richness  = m.gb.rich.reb,
  GB_Diversity = m.gb.shan.reb
)

# Pull out model results to create a table

# Function to extract coefficients for lmrob or glmmTMB
extract_coefs <- function(model, response_name) {
  if (inherits(model, "lmrob")) {
    # tidy() works for lmrob
    tidy(model) %>%
      mutate(Model = response_name) %>%
      select(Model, term, estimate, std.error, statistic, p.value)
  } else if (inherits(model, "glmmTMB")) {
    # For glmmTMB, use tidy with effects="fixed"
    broom.mixed::tidy(model, effects = "fixed", conf.int = FALSE) %>%
      mutate(Model = response_name) %>%
      select(Model, term, estimate, std.error, statistic, p.value)
  } else {
    stop("Unknown model type")
  }
}

# Apply the function to all models
reb.coef_table <- imap_dfr(reb.models, ~extract_coefs(.x, .y)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 3)))  # round numbers for readability

# View the table
reb.coef_table

reb.model_table <- reb.coef_table
write.csv(reb.model_table, "Tables/Reburn_Effects.csv", row.names = FALSE)

```



```{r}



```



