---
output:
  pdf_document: default
  html_document: default
---

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(gtools)
library(vegan)
library(lubridate)
library(ggpubr)
library(pscl)
library(glmmTMB)
library(zoo)
library(rcompanion)
library(Rmisc)
library(MuMIn)
```

```{r}
lytle_rodents <- read.csv("Lytle_16_to_23_rodents.csv")
lytle_rodents <-  lytle_rodents %>% select(c(-EAR..mm.,-HINDFOOT..mm.,-TAIL..mm.,-Comments, -AGE, -RECAPTURE,-OBSERVERS,-TIME..24hr.,-PREGNANT,-CONDITION))
```

```{r}


lytle_rodents[lytle_rodents$SPECIES == "cactus mouse?",4] <- "PEER"
lytle_rodents[lytle_rodents$SPECIES == "DEIM",4] <- "DIME"
lytle_rodents[lytle_rodents$SPECIES == "DIEM",4] <- "DIME"
lytle_rodents[lytle_rodents$SPECIES == "KRAT",4] <- "DIME"
lytle_rodents[lytle_rodents$SPECIES == "Chipmunk",4] <- "AMLE"
lytle_rodents[lytle_rodents$SPECIES == "grasshopper",4] <- "ONTO"
lytle_rodents[lytle_rodents$SPECIES == "grasshopper mouse",4] <- "ONTO"
lytle_rodents[lytle_rodents$SPECIES == "grasshopper mouse, ONTO",4] <- "ONTO"
lytle_rodents[lytle_rodents$SPECIES == "grasshopper mouse?",4] <- "ONTO"
lytle_rodents[lytle_rodents$SPECIES == "grasshopper mouse? ONTO?",4] <- "ONTO"
lytle_rodents[lytle_rodents$SPECIES == "grasshopper or cactus mouse",4] <- "ONTO"
lytle_rodents[lytle_rodents$SPECIES == "Woodrat",4] <- "NELE"
lytle_rodents[lytle_rodents$SPECIES == "NELEL",4] <- "NELE"
lytle_rodents[lytle_rodents$SPECIES == "PEMA/PECR",4] <- "PEMA"


lytle_rodents <- subset(lytle_rodents, SPECIES != "lizard")
lytle_rodents <- subset(lytle_rodents, SPECIES != "Jack Rabbit")
lytle_rodents <- subset(lytle_rodents, SPECIES != "X")
lytle_rodents <- subset(lytle_rodents, SPECIES != "cottontail")
lytle_rodents <- subset(lytle_rodents, SPECIES != "PEMA/CHFO")

lytle_rodents$EVENT <- format(as.Date(lytle_rodents$DATE, format="%m/%d/%Y"),"%m/%Y")

lytle_rodents$DATE <- as.Date(lytle_rodents$DATE, format = "%m/%d/%Y")


```

Create a dataframe of all possible combinations

```{r}

# Define the column values
years <- 2016:2023
months <- c(4, 5, 7, 10, 11)
rodents <- c("N", "S", "C")
burned <- c("B", "U", "C")
trap_days <- 1:3
blocks <- 1:5


# Create all possible combinations
combinations <- expand.grid(Year = years, Month = months, Rodents = rodents,
                            Burned = burned, `Trap Day` = trap_days, Block = blocks)

combinations<- filter(combinations, Year < 2022 | `Trap Day` < 3)

combinations <- subset(combinations, !(Year == 2016 & !(Month %in% c(4, 7))))
combinations <- subset(combinations, !(Year == 2017 & !(Month %in% c(4, 7, 10))))
combinations <- subset(combinations, !(Year == 2018 & !(Month %in% c(4, 7, 10))))
combinations <- subset(combinations, !(Year == 2019 & !(Month %in% c(5, 11))))
combinations <- subset(combinations, !(Year == 2020 & !(Month %in% c(4, 7, 10))))
combinations <- subset(combinations, !(Year == 2021 & !(Month %in% c(4, 10))))
combinations <- subset(combinations, !(Year == 2022 & !(Month %in% c(4))))
combinations <- subset(combinations, !(Year == 2023 & !(Month %in% c(4))))

combinations <- filter(combinations, Rodents !=  "C" | Burned != "B")
combinations <- filter(combinations, Rodents !=  "C" | Burned != "U")

combinations <- filter(combinations, Rodents !=  "N" | Burned != "C")
combinations <- filter(combinations, Rodents !=  "S" | Burned != "C")

combinations$code <- paste(combinations$Year, combinations$Month, combinations$Rodents, combinations$Burned, combinations$Block, combinations$`Trap Day`)

```


This is to create the usable file that takes into account all trapdays including blanks and only using uniqely caught individuals from each trapping event. This is a conservative estimate of abundance

```{r}

#### Create a data file with population estimates per trap event

df <- lytle_rodents

# I need a recapture columns

#### I ask ChatGPT how to find recaptures and this is what it gave me and it works
df <- df %>% 
  dplyr::group_by(EVENT, SPECIES, EARTAG..) %>% 
  dplyr::mutate(Recapture = ifelse(row_number() == 1, "C", "R")) %>% 
  ungroup()

df$Recapture <-as.factor(df$Recapture)


#### This is to make it only uniquely caught individuals
### To get total captures, we can skip this 
df <- df[df$Recapture== "C",]

#head(df)
#head(combinations)

#format so each species is a column with total counts per trapping event
#we lose mass and sex
df <- df %>% 
  dplyr::count(SPECIES, EVENT, BURNED, RODENTS, TRAP.DAY, DATE, BLOCK, PLOT)%>%
  pivot_wider(names_from = SPECIES, values_from = n) 

df <-na.replace(df, 0) # Replace NA's with 0's
df$DATE <- as.Date(df$DATE, format = "%m/%d/%Y")
df$BURNED <- as.factor(df$BURNED)
df$RODENTS <- as.factor(df$RODENTS)

df$code <- paste(year(df$DATE), month(df$DATE), df$RODENTS, df$BURNED, df$BLOCK, df$TRAP.DAY)

#summary(df)

df <- merge(combinations, df, by = "code", all.x = TRUE)

df<- df %>% select(c(-EVENT, -BURNED, -RODENTS, -TRAP.DAY, -BLOCK, -PLOT))

### I have NA's for certain dates and this is how chatGPT said to fill em in and it works


# Iterate over each row of the dataframe
for (i in 1:nrow(df)) {
  # Check if 'date' is NA
  if (is.na(df$DATE[i])) {
    # Find rows with the same 'year', 'month', and 'trap day'
    same_date_rows <- df$Year == df$Year[i] &
      df$Month == df$Month[i] &
      df$`Trap Day` == df$`Trap Day`[i]
    # Get the first non-NA value of 'date' from the matching rows
    replacement_date <- na.omit(df$DATE[same_date_rows])[1]
    
    # Replace NA with the 'replacement_date'
    df$DATE[i] <- replacement_date
  }
}
 #Now go back and re run na.replace

#### Get rid of trap days

df$`Trap Day` <- as.factor(df$`Trap Day`)
df$Block <- as.factor(df$Block)

##First add climate

lytle_clim <- read.csv("C:/Users/ryanp/Documents/Dissertation/Raw data/Environmental Spatial Data/Climate/PRISM_data_Lytle.csv")
head(lytle_clim)

 #ggplot(data = lytle_clim, aes(Date, ppt..mm.))+
#  geom_point()+
#  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))

lytle_clim$Date <- as.Date(paste(lytle_clim$Date,01,sep = "-"))
lytle_clim$EVENT <- format(lytle_clim$Date, "%m/%Y")

lytle_clim$Year <- lubridate::year(lytle_clim$Date)
lytle_clim$Month <- month(lytle_clim$Date)

df <- merge(df, lytle_clim, all.x = TRUE, by = c("Year", "Month"))

df<- df %>% select(c(-Date))
df <- na.replace(df,0)
write.csv(df,"lytle_rodents_16to23.csv")

#Trap Day out
df <- df %>%
 dplyr::group_by(Year, Month, Burned, Rodents, Block) %>%
  dplyr::summarise(across(4:18,sum), across(19:25, mean))

df <- na.replace(df, 0)


```




Check if the number of rodents changes over time



```{r}

rodents_time <- ggplot(data = lytle_rodents, aes(DATE, fill = SPECIES))+
  geom_bar(width = 50)+
  theme(axis.text.x = element_text(angle=90, vjust=.5, hjust=1))

rodents_time

```

calculate richness and abundance functions

```{r}

calc_rich <- function(data, col_start, col_end) {
  columns <- c(col_start:col_end)
  subset_data <- data[, columns]
  positive_columns_per_row <- rowSums(subset_data > 0)
  return(positive_columns_per_row)
}

calc_abun <- function(data, col_start, col_end) {
  columns <- c(col_start:col_end)
  subset_data <- data[, columns]
  total <- rowSums(subset_data)
  return(total)
}

```

calculate richness, shannons diversity, simpsons diversity and abundance


```{r}

df$RICHNESS <- calc_rich(df,6,20)
df$SH_DIVERSITY <- diversity(df[,6:20], index = "shannon")
df$SI_DIVERSITY <- diversity(df[,6:20], index = "simpson")
df$ABUNDANCE <- calc_abun(df,6,20)


```


Histograms of metrics

```{r}

hist(df$RICHNESS)
hist(df$SH_DIVERSITY)
hist(df$SI_DIVERSITY)
hist(df$ABUNDANCE)

```




make some graphs

```{r}

attach(df)

ggabun <- ggplot(data = df[df$Burned == c("B", "U"),],  aes(Burned, ABUNDANCE, fill = Rodents))+
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "mean") +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.2)

ggabun

ggrich <- ggplot(data = df[df$Burned == c("B", "U"),],  aes(Burned, RICHNESS, fill = Rodents))+
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "mean") +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.2)

ggrich

ggsi <- ggplot(data = df[df$Burned == c("B", "U"),],  aes(Burned, SI_DIVERSITY, fill = Rodents))+
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "mean") +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.2)

ggsi

ggsh <- ggplot(data = df[df$Burned == c("B", "U"),],  aes(Burned, SH_DIVERSITY, fill = Rodents))+
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "mean") +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(0.9), width = 0.2)

ggsh




```

Plot the effectiveness of the rodent exclusion over time First have to calculate time since fire to give a numeric x-axis for the regression

```{r}


df$TSF <- as.numeric(df$Year-2011)

```

Now plot the regression between abundance in Rodent Exclusion plots and time since fire

```{r}



ggplot(data = df[df$Rodents=="N",], aes(x=TSF, y=ABUNDANCE))+
  geom_point()+
  geom_smooth(method = "lm")+
  stat_regline_equation(label.y = c(8,10), aes(label = after_stat(eq.label))) +
  stat_regline_equation(label.y = c(7,6), aes(label = after_stat(rr.label)))

```

Test assumptions for linear regression use log for abundance and richness because its one-inflated instead of zero inflated and that corrects it The data is not normal so we're going to go with a Kruskal Wallis instead of an ANOVA, I'm leaving the ANOVA code up above for in case but we will only report on the Kruskal Wallis

\*Update Kruskal Wallis can't handle multiple explanatory variables so I'm switching to Scheiere Ray Hare from the rcompanion package



Data appears to be zero-inflated That means I'm going to have to use a zero inflated mixed effects model from glmmtmb when accounting for climate data, here is the Scheirer Ray Hare to test the two treatments

```{r}


srh_rich<- scheirerRayHare(RICHNESS ~ Burned + Rodents + Burned*Rodents, data = df)
srh_rich

srh_sh <- scheirerRayHare(data = df, SH_DIVERSITY ~ Burned + Rodents + Burned*Rodents) 
srh_sh

srh_si <- scheirerRayHare(data = df, SI_DIVERSITY ~ Burned + Rodents + Burned*Rodents)
srh_si

srh_abun <- scheirerRayHare(data = df, ABUNDANCE ~ Burned + Rodents + Burned*Rodents)
srh_abun

```




Import climate data for the climate model

I've added annual precip from the scrap code

```{r}

pptrich <- ggplot(data = df, aes(ppt..mm., RICHNESS, color = Burned))+
  geom_point()+ geom_smooth(method = "lm")+
  #stat_regline_equation(aes(label = after_stat(eq.label))) +
  stat_regline_equation(aes(label = after_stat(rr.label)))

pptrich+geom_jitter()

tmrich <- ggplot(data = df, aes(tmean..degrees.C., RICHNESS, color = Burned))+
  geom_point()+ geom_smooth(method = "lm")+
  #stat_regline_equation(aes(label = after_stat(eq.label))) +
  stat_regline_equation(aes(label = after_stat(rr.label)))

tmrich+geom_jitter()

```

So now I've added the climate data for the abundance of each species per treatment, now I can start testing models

First though I have to check for autocorrelated variables because I forgot that. SO I'm going to make a correlation matrix and then remove variables that have a correlation coefficient higher than 0.6.

I get NA's I'm not sure why so thats something I better figure out

```{r}

cor_matrix_pe <- as.matrix(lytle_clim[,c(2:8)])

head(cor_matrix_pe)
cor(cor_matrix_pe)



```
Looks like I only need precip and mean temp

Let me plot them first

```{r}

df$Date <- paste(df$Year,df$Month, 01, sep = "-")
df$Date <- as.Date(df$Date, format = "%Y-%m-%d")
df$Plot <- paste(df$Burned,df$Rodents, sep = "")

ppt_plot <- ggplot(data = df, aes(x = Date, y = ppt..mm.)) +
  geom_line() +
  labs(x = "Date", y = "Precipitation by month (mm)") +
  scale_color_discrete(name = "Category")

ppt_plot

tmean_plot <- ggplot(data = df, aes(x = Date, y = tmean..degrees.C.)) +
  geom_line() +
  labs(x = "Date", y = "mean monthly temperature") +
  scale_color_discrete(name = "Category")

tmean_plot


```



Now I'm going to try with my df_p (Richness, Abundance and Diversity), Note this doesn't include mark recapture data yet. All captures are counted in abundance including recaptures, I may go back and fix it




More graphs: I want to show richness, abundance and diversity over time

```{r}

#I'm going to have to get an average of each block 
df$Date <- paste(df$Year,df$Month, 01, sep = "-")
df$Date <- as.Date(df$Date, format = "%Y-%m-%d")
df$Plot <- paste(df$Burned,df$Rodents, sep = "")


pd <- position_dodge(0.1) # move them .05 to the left and right

df_rich <- summarySE(df, 
                        measurevar = c("RICHNESS"),
                        groupvars = c("Plot", "Date"))

rich_time <- ggplot(data = df_rich, aes(x = Date, y = RICHNESS, group = Plot)) +
  geom_line(size = 0.5) +
  geom_point(size = 3, aes(shape = Plot)) +
  labs(x = "Date", y = "Richness") +
  scale_shape_manual(values = c(15, 17, 8, 0, 2)) +
  theme_pubclean()



df_abun <- summarySE(df, 
                        measurevar = c("ABUNDANCE"),
                        groupvars = c("Plot", "Date"))

abun_time <- ggplot(data = df_abun, aes(x = Date, y = ABUNDANCE, group = Plot)) +
  geom_line(size = 0.5) +
  geom_point(size = 3, aes(shape = Plot)) +
  labs(x = "Date", y = "Abundance") +
  scale_shape_manual(values = c(15, 17, 8, 0, 2)) +
  theme_pubclean()



df_sh <- summarySE(df, 
                        measurevar = c("SH_DIVERSITY"),
                        groupvars = c("Plot", "Date"))

sh_time <- ggplot(data = df_sh, aes(x = Date, y = SH_DIVERSITY, group = Plot)) +
  geom_line(size = 0.5) +
  geom_point(size = 3, aes(shape = Plot)) +
  labs(x = "Date", y = "Shannon's Diversity") +
  scale_shape_manual(values = c(15, 17, 8, 0, 2)) +
  theme_pubclean()

df_si <- summarySE(df, 
                        measurevar = c("SI_DIVERSITY"),
                        groupvars = c("Plot", "Date"))

si_time <- ggplot(data = df_si, aes(x = Date, y = SI_DIVERSITY, group = Plot)) +
  geom_line(size = 0.5) +
  geom_point(size = 3, aes(shape = Plot)) +
  labs(x = "Date", y = "Simpsons Diversity") +
  scale_shape_manual(values = c(15, 17, 8, 0, 2)) +
  theme_pubclean()
  
rich_time
abun_time
sh_time
si_time


```

ANOVA's
```{r}

#### Richness ####

# zero inflated response variable

str(df)

rich_zin <- aov(data = df, RICHNESS ~  Rodents + Burned + Rodents*Burned + (1|Block))

summary(rich_zin)

```



Models from the graphs

An idea for solving the zero inflation is to do separate analyses on 0 vs >1 that is just counted as 1 and the second analysis without zeros and every number greater than 1 

I need to add all the zeros that were already there though and that breaks it

so this doesnt work so ill just carry on with my zero inflated model



Add the cumulative precipitation and lag effects

```{r}

##### Create different combinations of rainfall data

combinations$Date <- paste(combinations$Year,combinations$Month,"01", sep = "-")
clim <- lytle_clim


#ChatGPT to calculate previous 3 months precipitation
# Iterate over each row in the dataframe
i=0
for (i in 1:nrow(clim)) {
  # Check if there are at least three previous rows
  if (i >= 4) {
    # Calculate the sum of precipitation for the previous three months
    prev_3_months_sum <- sum(clim$ppt..mm.[(i-2):i-1])
    # Assign the sum to the new column in the current row
    clim$pre3[i] <- prev_3_months_sum
  }
}

### 6 Months ###
i=0
clim$pre6 <- NA
for (i in 1:nrow(clim)) {
  # Check if there are at least three previous rows
  if (i >= 6) {
    # Calculate the sum of precipitation for the previous three months
    prev_3_months_sum <- sum(clim$ppt..mm.[(i-5):i-1])
    # Assign the sum to the new column in the current row
    clim$pre6[i] <- prev_3_months_sum
  }
}


### 12 Months ###
i=0
clim$pre12 <- NA
for (i in 1:nrow(clim)) {
  # Check if there are at least three previous rows
  if (i >= 12) {
    # Calculate the sum of precipitation for the previous three months
    prev_3_months_sum <- sum(clim$ppt..mm.[(i-11):i-1])
    # Assign the sum to the new column in the current row
    clim$pre12[i] <- prev_3_months_sum
  }
}


##### 1 Year lag ####
i=0
clim$lag1 <- NA
for (i in 1:nrow(clim)) {
  # Check if there are at least three previous rows
  if (i >= 13) {
    # Calculate the sum of precipitation for the previous three months
    lag <- clim$pre12[i-1]
    # Assign the sum to the new column in the current row
    clim$lag1[i] <- lag
  }
}

clim <- clim %>% select(Date,pre3,pre6,pre12,lag1)

df <- merge(df, clim, all.x = TRUE, by = "Date")


```


Time to start building some actual models

```{r}

#### I'm going to start by doing abundance for all of candidate models

df <- df %>% select(Date, Burned, Rodents, Block, ABUNDANCE, RICHNESS, SH_DIVERSITY, SI_DIVERSITY,
                    ppt..mm., tmean..degrees.C., pre3, pre6, pre12, lag1, TSF)

abun1 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block), 
                 ziformula = ~1,
                 family = poisson)

abun2 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block)+Burned, 
                 ziformula = ~1,
                 family = poisson)

abun3 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block) + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = poisson)

abun4 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block) + Burned + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = poisson)

abun5 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block) + Burned + pre12, 
                 ziformula = ~1,
                 family = poisson)

abun6 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block) + Burned + pre6, 
                 ziformula = ~1,
                 family = poisson)

abun7 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block) + Burned + lag1, 
                 ziformula = ~1,
                 family = poisson)

abun8 <- glmmTMB(data = df, ABUNDANCE ~ (1|Block) + Burned + lag1 + TSF, 
                 ziformula = ~1,
                 family = poisson)



#df.ex <- na.exclude(df)

#abun_glob <- glmmTMB(data = df.ex, ABUNDANCE ~ (1|Block) + Burned + lag1 + ppt..mm.+tmean..degrees.C.+
#                       pre3+ pre6 + pre12,  
#                 ziformula = ~1,
#                 family = poisson, na.action = "na.fail")

#dredge(abun_glob)

summary(abun8)


model.sel(abun1,abun2, abun3,abun4,abun5,abun6,abun7)

#### Richness

rich1 <- glmmTMB(data = df, RICHNESS ~ (1|Block), 
                 ziformula = ~1,
                 family = poisson)

rich2 <- glmmTMB(data = df, RICHNESS ~ (1|Block)+Burned, 
                 ziformula = ~1,
                 family = poisson)

rich3 <- glmmTMB(data = df, RICHNESS ~ (1|Block) + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = poisson)

rich4 <- glmmTMB(data = df, RICHNESS ~ (1|Block) + Burned + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = poisson)

rich5 <- glmmTMB(data = df, RICHNESS ~ (1|Block) + Burned + pre12, 
                 ziformula = ~1,
                 family = poisson)

rich6 <- glmmTMB(data = df, RICHNESS ~ (1|Block) + Burned + pre6, 
                 ziformula = ~1,
                 family = poisson)

rich7 <- glmmTMB(data = df, RICHNESS ~ (1|Block) + Burned + lag1, 
                 ziformula = ~1,
                 family = poisson)

summary(rich7)

model.sel(rich1, rich2, rich3, rich4, rich5, rich6, rich7)

### SI_DIVERSITY
library(mgcv)
si1 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block), 
                 ziformula = ~1,
                 family = poisson)

si2 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block)+Burned, 
                 ziformula = ~1,
                 family = poisson)

si3 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block) + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = poisson)

si4 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block) + Burned + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = poisson)

si5 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block) + Burned + pre12, 
                 ziformula = ~1,
                 family = poisson)

si6 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block) + Burned + pre6, 
                 ziformula = ~1,
                 family = poisson)

si7 <- glmmTMB(data = df, SI_DIVERSITY ~ (1|Block) + Burned + lag1, 
                 ziformula = ~1,
                 family = poisson)

summary(si1)

sh1 <- glmmTMB(data = df, SH_DIVERSITY ~ (1|Block), 
                 ziformula = ~1,
                 family = "gaussian")

sh2 <- glmmTMB(data = df, SH_DIVERSITY+1 ~ (1|Block)+Burned, 
                 ziformula = ~1,
                 family = "gaussian")


sh3 <- glmmTMB(data = df, SH_DIVERSITY ~ (1|Block) + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = "gaussian")

sh4 <- glmmTMB(data = df, SH_DIVERSITY ~ (1|Block) + Burned + ppt..mm.+tmean..degrees.C., 
                 ziformula = ~1,
                 family = "gaussian")

sh5 <- glmmTMB(data = df, SH_DIVERSITY  ~  (1|Block) + Burned + pre12, 
                 ziformula = ~1,
                 family = "gaussian")

sh6 <- glmmTMB(data = df, SH_DIVERSITY ~ (1|Block) + Burned + pre6, 
                 ziformula = ~1,
                 family = "gaussian")

sh7 <- glmmTMB(data = df, SH_DIVERSITY ~ (1|Block) + Burned + lag1, 
                 ziformula = ~1,
                 family = "gaussian")

model.sel(sh1,sh2,sh3,sh4,sh5,sh6,sh7)

summary(sh5)



```



